### 2016-05-02 paper/paper.tex
paper full search

\documentclass[runningheads,a4paper]{llncs}

\usepackage{amssymb}
\setcounter{tocdepth}{3}
\usepackage{graphicx}
\usepackage{pgfplots}

\usepackage{url}
\newcommand{\keywords}[1]{\par\addvspace\baselineskip
\noindent\keywordname\enspace\ignorespaces#1}

\begin{document}

\mainmatter

\title{Fast Nearest Neighbour Classification}
\subtitle{\textnormal{\small{Seminar\\
Intelligente Software-Systeme\\
Sommersemester 2015\\\vspace{1\baselineskip}
Stefan Fricke\\\vspace{2\baselineskip}
April 2016\\
Betreuer: Stephan Spiegel\\\vspace{1\baselineskip}}}}

\titlerunning{Fast Nearest Neighbour Classification}

\author{Gordon Lesti\\313249\\Studiengang: Informatik\\gordon.lesti@campus.tu-berlin.de\\\vspace{5\baselineskip}}

\authorrunning{Fast Nearest Neighbour Classification}

\institute{Technische Universit\"at Berlin\\
Fakult\"at IV Elektrotechnik und Informatik\\
Fachgebiet AOT\\
Prof. Dr. Sahin Albayrak\\
\url{http://www.aot.tu-berlin.de/}}

\toctitle{Fast Nearest Neighbour Classification}
\tocauthor{Gordon Lesti}
\maketitle

\begin{abstract}
This term paper is about nearest neighbour classifiers that work with the triangle inequality of metric spaces.
Benchmarks on examples from the two-dimensional space will show the advantages and disadvantages of three known
algorithms. The Orchard’s Algorithm, the Annulus Method and the AESA will be explained on examples and compared to the
full search. The benchmarks aren't comparing execution time, they will compare calls of the distance function.
\keywords{Nearest neighbour classifiers, Triangle inequality, Metric space}
\end{abstract}

\tableofcontents

\newpage

\section{Introduction}

\subsection{Requirements}

\section{Algorithms}

\subsection{Full Search}

The full search isn't a real fast nearest neighbour algorithm, it's more the default algorithm for such a problem. We
just iterate over all $x \in S$ and calculate $d(q, x)$. After that we just take $a \in S$, so that
$d(q, a) \le d(q, x)$ for all $x \in S$. The big advantage compared to the following algorithms is, that the full search
works also in not metric spaces. The full search has no preprocessing.

\subsection{Orchard’s Algorithm}

\subsubsection{Preprocessing}

\subsection{Annulus Method}

\subsubsection{Preprocessing}

\subsection{AESA}

\subsubsection{Preprocessing}

\section{Benchmarks}

\subsection{Implementation}

\subsection{Full Search}

\subsection{Orchard’s Algorithm}

\subsection{Annulus Method}

\subsection{AESA}

\subsection{Outcome}

\begin{center}
	\begin{tikzpicture}
		\begin{axis}[ymin=0, xmin=0, legend pos=north west]
			\addplot table {data/mainprocessing/fullsearch.csv};
            \addlegendentry{FullSearch}
            \addplot table {data/mainprocessing/orchards.csv};
            \addlegendentry{Orchards}
            \addplot table {data/mainprocessing/annulus.csv};
            \addlegendentry{Annulus}
            \addplot table {data/mainprocessing/aesa.csv};
            \addlegendentry{AESA}
		\end{axis}
	\end{tikzpicture}
\end{center}

\end{document}