### 2016-05-10 paper/paper.tex
something like ready

\documentclass[runningheads,a4paper]{llncs}

\usepackage{amssymb}
\setcounter{tocdepth}{3}
\usepackage{graphicx}
\usepackage{pgfplots}
\usepackage{algorithm}
\usepackage{algorithmic}

\usepackage{url}
\newcommand{\keywords}[1]{\par\addvspace\baselineskip
\noindent\keywordname\enspace\ignorespaces#1}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}

\begin{document}

\mainmatter

\title{Fast Nearest Neighbour Classification}
\subtitle{\textnormal{\small{Seminar\\
Intelligente Software-Systeme\\
Sommersemester 2015\\\vspace{1\baselineskip}
Stefan Fricke\\\vspace{2\baselineskip}
April 2016\\
Betreuer: Stephan Spiegel\\\vspace{1\baselineskip}}}}

\titlerunning{Fast Nearest Neighbour Classification}

\author{Gordon Lesti\\313249\\Studiengang: Informatik\\gordon.lesti@campus.tu-berlin.de\\\vspace{5\baselineskip}}

\authorrunning{Fast Nearest Neighbour Classification}

\institute{Technische Universit\"at Berlin\\
Fakult\"at IV Elektrotechnik und Informatik\\
Fachgebiet AOT\\
Prof. Dr. Sahin Albayrak\\
\url{http://www.aot.tu-berlin.de/}}

\toctitle{Fast Nearest Neighbour Classification}
\tocauthor{Gordon Lesti}
\maketitle
\pagenumbering{roman}

\addcontentsline{toc}{section}{Abstract}
\begin{abstract}
This term paper is about nearest neighbour classifiers which work with the triangle inequality of metric spaces. The
Orchard’s Algorithm, the Annulus Method and the AESA will be explained on examples and compared to the Full Search.
Benchmarks will show the advantages and disadvantages of the three algorithms. Counting the calls of the distance
function is the focus of the benchmarks.
\keywords{Nearest neighbour classifiers, Triangle inequality, Metric space}
\end{abstract}

\tableofcontents
\addcontentsline{toc}{section}{Table of Contents}

\newpage
\pagenumbering{arabic}

\section{Introduction}

HERE COMES THE INTRODUCTION.

\subsection{Description of the problem}

The mission of a nearest neighbour search algorithm is to find in a given set of items, that item with the smallest
distance to a given query item.\\
The requirements are:
\begin{itemize}
	\item a set $\mathbb{U}$
	\item a distance function $D$ on $\mathbb{U}$, with $d: \mathbb{U} \times \mathbb{U} \to \mathbb{R}$
	\item a set $S \subset \mathbb{U}$ with size $n$
	\item a query $q \in \mathbb{U}$
\end{itemize}
The solution is:
\begin{itemize}
	\item an $a \in S$, with $D(q, a) \le D(q, x)$ for all $x \in S$
\end{itemize}

\section{Algorithms}

Every nearest neighbour search algorithm can be divided into two steps.
\paragraph{Preprocessing}
During the preprocessing the algorithm prepares the given set $\mathbb{S}$ for his needs.

\paragraph{Query Processing}
The query processing is the main nearest neighbour search, the algorithm takes the query $q$ as input and works on the
preprocessed data to find the nearest neighbour $a$.\\

The following section will explain the preprocessing and query processing of the Full Search, the Orchard’s Algorithm,
the Annulus Method and the AESA.

\subsection{Full Search}

The Full Search is the upper bound for the following algorithms. Exactly $n$ calls of the distance function are required
to find the nearest neighbour in a given $S$ of size $n$.\\

\begin{algorithm}[H]
	\caption{Full Search}
	\label{fullsearch}
	\begin{algorithmic}
		\REQUIRE query $q \in \mathbb{U}$, set $S \subset \mathbb{U}$ of size $n$, distance function $D$
		\ENSURE $a \in S$ with $D(q, a) \le D(q, x)$ $\forall x \in S$
		\STATE $r := \infty$
		\STATE $a :=$ NULL
		\FORALL{$x \in S$}
			\STATE $xd := D(x, q)$
			\IF{$sd < r$}
				\STATE $r := xd$
				\STATE $a := x$
			\ENDIF
		\ENDFOR
		\RETURN $a$
	\end{algorithmic}
\end{algorithm}

The advantage compared to the following algorithms is, that the Full Search works also in not metric spaces. The Full
Search needs no preprocessing.

\subsection{Orchard’s Algorithm}

\subsubsection{Preprocessing}

The preprocessing of the Orchard’s Algorithm has a high complexity. For every $p \in S$, the algorithm will generate a
list that contains all $x \in S\setminus\left\{ {p}\right\}$ with their distance $D(p, x)$. Those lists will be ordered
ascending to the distance. For a set $S$ with size $n$, the preprocessing will call the distance function
$\frac{n(n-1)}{2}$ times and will sort $n$ lists of size $n-1$.

\subsubsection{Query Processing}

When calling the query processing with query $q$, the algorithm will randomly select one object $c \in S$ as initial
candidate and calculate $D(c, q)$. Afterwards the algorithm walks over the ordered list of $c$ and with every picked $s$
from the list the algorithm calculates $D(l, q)$. If $D(l, q)$ is smaller than $D(c, q)$, $l$ is the new candidate $c$
and the algorithm walks over the list of $l$. The algorithm aborts if it reaches the end of a list or if
$D(c, s) > 2D(c, q)$. The current candidate $c$ is the nearest neighbour.

\subsection{Annulus Method}

The Annulus Method works a little bit similar to the Orchard’s Algorithm, but with a less complex preprocessing.

\subsubsection{Preprocessing}

Instead of generating a list for all $x \in S$, the Annulus Method generates just one list for a random $p^* \in S$. In
the preprocessing the algorithm generates one list that contains all $x \in S$ with their distance $D(p^*, x)$. This
list will be ordered ascending to the distance. For a set $S$ with size $n$, the preprocessing will call the distance
function $n$ times and will sort one list of size $n$.

\subsubsection{Query Processing}

The query processing of the Annulus Method starts by picking a random candidate $c \in S$ from the ordered list of
$p^*$. Afterwards the algorithm walks alternating away from $p^*$ and back to him in the list. If the current object $s$
has $D(s, q) < D(c, q)$, the algorithm sets $s$ as the new candidate $c$. If the current object $s$ is under $c$ in the
list and $D(p^*, s) < D(p^*, q) - D(c, q)$, the algorithm can ignore all objects under $s$ in the list. If the current
object $s$ is above $c$ in the list and $D(p^*, s) > D(p^*, q) + D(c, q)$, the algorithm can ignore all objects above
$s$ in the list. The object $c$ is the nearest neighbour, if the entire list is traversed.

\subsection{AESA}

\subsubsection{Preprocessing}

The preprocessing for the AESA is starting the same way as the preprocessing of the Orchard’s Algorithm. The algorithm
calculates the distance $D(x, y)$ of every $x \in S$ to every $y \in S\setminus\left\{ {x}\right\}$. But instead of
writing those distances into lists and sorting them, the algorithm writes those distances into a matrix to access them
quickly. For a set $S$ with size $n$, the preprocessing will call the distance function $\frac{n(n-1)}{2}$ times and
will sort $n$ lists of size $n-1$.

\subsubsection{Query Processing}

When starting with the query processing, the algorithm generates a list for all objects $x \in S$ with a bounded
distance $D_P(x, q) = -\infty$ and initialize $r = \infty$ as smallest distance to $q$. Afterwards the algorithm
iterates the following steps until the list is empty.
\begin{itemize}
	\item pick and remove $x$ from the list with the smallest $D_p(x, q)$ or a random one in the first iteration
	\item calculate $D(x, q)$
	\item set $r = D(x, q)$ if $r > D(x, q)$ and remember $x$ as current nearest neighbour
	\item for all $y$ from the list, set $D_P(y, q) = max(D_P(y, q), |D(x, q) - D(x, y)|)$ and remove $y$ from the list
		if the new $D_P(y, q) > r$
\end{itemize}

\section{Benchmarks}

As described above, the following benchmarks will not compare the runtime of the mentioned algorithms. The benchmarks
will compare the calls of the distance function that the algorithm needs for preprocessing and for finding a nearest
neighbour. Those benchmarks can make sense when the distance function has a heavy complexity. For example the string
metric Levenshtein distance or in higher dimensional vector spaces.\\
In general a benchmark is affected by the implementation, by the executing machine, by the domain of the problem
and the distribution of the example. As the following benchmarks only counting the calls of the distance function, they
aren't affected by the implementation or the executing machine.

\subsection{Implementation}

The algorithms above are all implemented in Java and can be found in the following git repository\\
\url{https://github.com/GordonLesti/FastNearestNeighbourClassification}.\\

A nearest neighbour algorithm implemented in Java should handle every type of object that a set $\mathbb{U}$ can
represent under the condition that there is a metric on $\mathbb{U}$. Because of that, a nearest neighbour algorithm
class needs to extend from the abstract class\\ \verb+FastNearestNeighbourClassificator<T, D extends Comparable<D>>+
that can be found in the package \verb+fnnc.algo+. The generic \verb+T+ represents an object from the set $\mathbb{U}$
and the generic \verb+D+ represents the distance and has to be of type \verb+Comparable+ from the package
\verb+java.lang+.\\

A possible example for the type \verb+T+ can be \verb+Point2D.Double+ from the package
\verb+java.awt.geom+ and a possible way to implement a distance function would be the euclidean distance of type
\verb+Double+.

\subsubsection{FastNearestNeighbourClassificator}

The abstract class\\ \verb+FastNearestNeighbourClassificator<T, D extends Comparable<D>>+ has two protected class
variables:
\begin{itemize}
	\item a \verb+distanceCalculator+ of type \verb+DistanceCalculator<T, D>+
	\item a \verb+objectCollection+ of type \verb+Collection<T>+
\end{itemize}
and three simple class methods:
\begin{itemize}
	\item a constructor that takes an object of type \verb+DistanceCalculator<T, D>+ from the package
		\verb+fnnc.model+ and sets the protected class variable
	\item a default function \verb+preProcessing+ that takes a \verb+Collection<T>+ from the package
		\verb+java.util+ and sets the protected class variable
	\item a abstract function \verb+calculateNearestNeighbour+ that takes the query object of type \verb+T+ and
		returns the nearest neighbour object of type \verb+T+
\end{itemize}

A instance of type \verb+FastNearestNeighbourClassificator+ shouldn't and can't calculate the distance between two
objects. So that a instance of \verb+DistanceCalculator+ should be injected into the constructor. All implemented
algorithms exclude the \verb+FullSearch+ require that the \verb+DistanceCalculator+ returns a distance of type
\verb+Double+ instead of only \verb+Comparable+. This is connected with the fact that only the Full Search can operate
in a not metric space.

\subsubsection{DistanceCalculator}

The abstract class\\ \verb+DistanceCalculator<Point2D.Double, Double>+ has just one private \verb+count+ of type
\verb+int+ and three class methods:
\begin{itemize}
	\item a public final function \verb+calculateDistance+ that takes two objects of type \verb+T+, increases the
		internal count and returns the result of \verb+internalCalculateDistance+
	\item a public final function \verb+reset+ that resets the internal count
	\item a protected abstract function \verb+internalCalculateDistance+ that takes two objects of type \verb+T+ and
		returns the distance object of type \verb+D+
\end{itemize}

\subsection{Example data}

The benchmarks will use a \verb+java.util.Collection+ of type\\ \verb+java.awt.geom.Point2D+ that are representing
points in $\mathbb{R}^2$. For every size $n \in \left\{ {100, 200, \dots, 1900, 2000}\right\}$, every algorithm will get
thousand sets of size $n$ with randomly generated points. The class\\ \verb+fnnc.model.point2ddouble.DistanceCalculator+
will count every call of the function \verb+internalCalculateDistance+ and after all the application will calculate the
average of the calls for every algorithm.

\subsection{Result}

\subsubsection{Preprocessing}

\begin{figure}
	\begin{center}
		\begin{tabular}{ | l | c | c | c | c | c |}
			\hline
		  	& Full Search & Orchard & Orchard MarkBits & Annulus & AESA \\ \hline
		  	$n$ & 0 & $\frac{n(n-1)}{2}$ & $\frac{n(n-1)}{2}$ & $n$ & $\frac{n(n-1)}{2}$ \\ \hline
		\end{tabular}
	\end{center}
	\caption{The calls of the distance function $D$ in the preprocessing depending on the size $n$ of the test data set
		$S$.}
	\label{fig:preprocessing:tabular}
\end{figure}

\begin{figure}
	\begin{center}
		\begin{tikzpicture}
			\begin{axis}[
				ymin=0,
				xmin=0,
				xmax=2000,
				xlabel=size of test data set $S$,
				ylabel=calls of distance function $D$,
				legend pos=outer north east]
				\addplot table {data/preprocessing/fullsearch.csv};
	            \addlegendentry{Full Search}
	            \addplot table {data/preprocessing/orchards.csv};
	            \addlegendentry{Orchard}
				\addplot table {data/preprocessing/orchardsmb.csv};
				\addlegendentry{Orchard MarkBits}
	            \addplot table {data/preprocessing/annulus.csv};
	            \addlegendentry{Annulus}
	            \addplot table {data/preprocessing/aesa.csv};
	            \addlegendentry{AESA}
			\end{axis}
		\end{tikzpicture}
	\end{center}
	\caption{The average calls of the distance function $D$ in the preprocessing depending on the size of the test data
		set $S$ visualized with graphs.}
	\label{fig:preprocessing:graph}
\end{figure}

\subsubsection{Query Processing}

\begin{figure}
	\begin{center}
		\begin{tabular}{ | l | c | c | c | c | c |}
			\hline
		  	& Full Search & Orchard & Orchard MarkBits & Annulus & AESA \\ \hline
		  	100 & 100 & 29 & 20 & 46 & 4 \\ \hline
			200 & 200 & 39 &26 & 82 & 4 \\ \hline
			300 & 300 & 46 & 30 & 117 & 4 \\ \hline
			400 & 400 & 53 & 35 & 160 & 4 \\ \hline
			500 & 500 & 57 & 40 & 189 & 4 \\ \hline
			600 & 600 & 63 & 42 & 228 & 4 \\ \hline
			700 & 700 & 68 & 44 & 263 & 4 \\ \hline
			800 & 800 & 73 & 46 & 303 & 4 \\ \hline
			900 & 900 & 78 & 50 & 341 & 4 \\ \hline
			1000 & 1000 & 79 & 51 & 367 & 4 \\ \hline
			1100 & 1100 & 84 & 56 & 409 & 4 \\ \hline
			1200 & 1200 & 86 & 56 & 442 & 4 \\ \hline
			1300 & 1300 & 89 & 59 & 466 & 4 \\ \hline
			1400 & 1400 & 90 & 60 & 510 & 4 \\ \hline
			1500 & 1500 & 97 & 63 & 541 & 4 \\ \hline
			1600 & 1600 & 100 & 66 & 573 & 4 \\ \hline
			1700 & 1700 & 102 & 67 & 624 & 4 \\ \hline
			1800 & 1800 & 104 & 69 & 660 & 4 \\ \hline
			1900 & 1900 & 109 & 71 & 664 & 4 \\ \hline
			2000 & 2000 & 113 & 73 & 740 & 4 \\ \hline
		\end{tabular}
	\end{center}
	\caption{The average calls of the distance function $D$ in the query processing depending on the size of the test
		data set $S$.}
	\label{fig:queryprocessing:tabular}
\end{figure}

\begin{figure}
	\begin{center}
		\begin{tikzpicture}
			\begin{axis}[
				ymin=0,
				xmin=0,
				xmax=2000,
				xlabel=size of test data set $S$,
				ylabel=average calls of distance function $D$,
				legend pos=outer north east]
				\addplot table {data/queryprocessing/fullsearch.csv};
	            \addlegendentry{Full Search}
	            \addplot table {data/queryprocessing/orchards.csv};
	            \addlegendentry{Orchard}
				\addplot table {data/queryprocessing/orchardsmb.csv};
				\addlegendentry{Orchard MarkBits}
	            \addplot table {data/queryprocessing/annulus.csv};
	            \addlegendentry{Annulus}
	            \addplot table {data/queryprocessing/aesa.csv};
	            \addlegendentry{AESA}
			\end{axis}
		\end{tikzpicture}
	\end{center}
	\caption{The average calls of the distance function $D$ in the query processing depending on the size of the test
		data set $S$ visualized with graphs.}
	\label{fig:queryprocessing:graph}
\end{figure}

\paragraph{Standard deviation}

As I saw that some of the algorithms needed more and more tries to stabilize the average result, I have calculated for
every size of a set and every algorithm the standard deviation. In general it is no big surprise that algorithms that
start with random initialization have fluctuating results, but especially the Annulus Method was very fragile.

\begin{figure}
	\begin{center}
		\begin{tabular}{ | l | c | c | c | c | c |}
			\hline
		  	& Full Search & Orchard & Orchard MarkBits & Annulus & AESA \\ \hline
		  	100 & 0 & 13 & 8 & 23 & 1 \\ \hline
			200 & 0 & 17 & 11 & 46 & 1 \\ \hline
			300 & 0 & 21 & 13 & 69 & 1 \\ \hline
			400 & 0 & 25 & 15 & 94 & 1 \\ \hline
			500 & 0 & 27 & 17 & 119 & 1 \\ \hline
			600 & 0 & 29 & 19 & 144 & 1 \\ \hline
			700 & 0 & 32 & 19 & 162 & 1 \\ \hline
			800 & 0 & 33 & 21 & 190 & 1 \\ \hline
			900 & 0 & 36 & 22 & 213 & 1 \\ \hline
			1000 & 0 & 37 & 23 & 232 & 1 \\ \hline
			1100 & 0 & 38 & 24 & 264 & 1 \\ \hline
			1200 & 0 & 40 & 26 & 277 & 1 \\ \hline
			1300 & 0 & 43 & 27 & 301 & 1 \\ \hline
			1400 & 0 & 43 & 27 & 339 & 1 \\ \hline
			1500 & 0 & 47 & 27 & 350 & 1 \\ \hline
			1600 & 0 & 46 & 29 & 374 & 1 \\ \hline
			1700 & 0 & 48 & 29 & 401 & 1 \\ \hline
			1800 & 0 & 49 & 31 & 429 & 1 \\ \hline
			1900 & 0 & 52 & 32 & 442 & 1 \\ \hline
			2000 & 0 & 52 & 33 & 487 & 1 \\ \hline
		\end{tabular}
	\end{center}
	\caption{The standard deviation of the calls of the distance function $D$ in the query processing depending on the
		size of the test data set $S$.}
	\label{fig:standarddeviation:tabular}
\end{figure}

\begin{figure}
	\begin{center}
		\begin{tikzpicture}
			\begin{axis}[
				ymin=0,
				xmin=0,
				xmax=2000,
				xlabel=size of test data set $S$,
				ylabel=standard deviation of calls of distance function $D$,
				legend pos=outer north east]
				\addplot table {data/standarddeviation/fullsearch.csv};
	            \addlegendentry{Full Search}
	            \addplot table {data/standarddeviation/orchards.csv};
	            \addlegendentry{Orchard}
				\addplot table {data/standarddeviation/orchardsmb.csv};
				\addlegendentry{Orchard MarkBits}
	            \addplot table {data/standarddeviation/annulus.csv};
	            \addlegendentry{Annulus}
	            \addplot table {data/standarddeviation/aesa.csv};
	            \addlegendentry{AESA}
			\end{axis}
		\end{tikzpicture}
	\end{center}
	\caption{The standard deviation of the calls of the distance function $D$ in the query processing depending on the
		size of the test data set $S$ visualized with graphs.}
	\label{fig:standarddeviation:graph}
\end{figure}

\end{document}