### 2016-06-22 presentation/presentation.tex
translate pres

\documentclass[handout]{beamer}

\usepackage[utf8]{inputenc}
\usepackage{default}
\usepackage{tikz}
\usepackage{color}

\begin{document}
\title{Fast Nearest Neighbour Classification}
\author{Gordon Lesti}
\date{July 17, 2015}

\frame{\titlepage}
\frame{\frametitle{Structure}\tableofcontents}

\section{Introduction}

\subsection{Problem}
\begin{frame}
 \frametitle{Nearest-Neighbour Searching}
 \begin{block}{Input}
   \begin{itemize}
    \item Set $\mathbb{U}$
    \pause
    \item Distance function $d$ on $\mathbb{U}$, with $d: \mathbb{U} \times \mathbb{U} \to \mathbb{R}$
    \pause
    \item Set $S \subset \mathbb{U}$ of size $n$
    \pause
    \item Query item $q \in \mathbb{U}$
  \end{itemize}
 \end{block}
 \pause
 \begin{block}{Output}
    \begin{itemize}
     \item Item $a \in S$, with $d(q, a) \le d(q, x)$ for all $x \in S$
    \end{itemize}
 \end{block}
\end{frame}

\subsection{Use}
\begin{frame}
 \frametitle{Use}
 \begin{itemize}
 \item Pattern recognition
 \item Statistical classification
 \item Image editing
 \item Coding theory
 \item Data compression
 \item Recommender system
 \item $\dots$
\end{itemize}
\end{frame}

\section{Solutions}

\subsection{Full Search}
\begin{frame}
 \frametitle{Full Search}
 \begin{itemize}
  \item Calculate $d(q, x)$ for all $x \in S$
  \pause
  \item Return $a \in S$, with $d(q, a) \le d(q, x)$ for all $x \in S$
 \end{itemize}
\end{frame}

% Beispiel1 Datenpunkte
\begin{frame}<handout:0>{Full Search}{Example}
 \begin{columns}[T]
  \begin{column}{.35\textwidth}
    $\mathbb{U} = \mathbb{R}^2$
    \begin{block}{Items}
      $x_1 = (3, 3)$\\
      $x_2 = (-1, 2)$\\
      $x_3 = (-4, -4)$\\
      $x_4 = (0, -1)$\\
      $x_5 = (4, -3)$
    \end{block}
  \end{column}
  \begin{column}{.6\textwidth}
     \begin{tikzpicture}[%
      scale=0.6,
      ]
      \clip (-5.0,-5.0) rectangle (5.0,5.0);
      \draw [help lines] (0,0) grid (6,5);
      \draw [help lines] (-6,0) grid (0,5);
      \draw [help lines] (0,-6) grid (6,0);
      \draw [help lines] (-6,-6) grid (0,0);
      \draw [color=black,->] (-5.0,0) -- (5.0,0);
      \draw [color=black,->] (0,-5.0) -- (0,5.0);

      \filldraw [color={rgb:red,0;green,0;blue,255}] (3,3) circle (2.5pt );
      \filldraw [color={rgb:red,0;green,0;blue,255}] (-1,2) circle (2.5pt );
      \filldraw [color={rgb:red,0;green,0;blue,255}] (-4,-4) circle (2.5pt );
      \filldraw [color={rgb:red,0;green,0;blue,255}] (0,-1) circle (2.5pt );
      \filldraw [color={rgb:red,0;green,0;blue,255}] (4,-3) circle (2.5pt );

      \node at (3.3,2.7) {$x_1$};
      \node at (-0.7,1.7) {$x_2$};
      \node at (-3.7,-4.3) {$x_3$};
      \node at (0.3,-1.3) {$x_4$};
      \node at (4.3,-3.3) {$x_5$};
      \end{tikzpicture}
  \end{column}
 \end{columns}
\end{frame}

% Beispiel1 Anfragepunkt
\begin{frame}{Full Search}{Example}
 \begin{columns}[T]
  \begin{column}{.35\textwidth}
    $\mathbb{U} = \mathbb{R}^2$
    \begin{block}{Items}
      $x_1 = (3, 3)$\\
      $x_2 = (-1, 2)$\\
      $x_3 = (-4, -4)$\\
      $x_4 = (0, -1)$\\
      $x_5 = (4, -3)$
    \end{block}
    \begin{block}{Query item}
      $q = (2, 1)$
    \end{block}
  \end{column}
  \begin{column}{.6\textwidth}
     \begin{tikzpicture}[%
      scale=0.6,
      ]
      \clip (-5.0,-5.0) rectangle (5.0,5.0);
      \draw [help lines] (0,0) grid (6,5);
      \draw [help lines] (-6,0) grid (0,5);
      \draw [help lines] (0,-6) grid (6,0);
      \draw [help lines] (-6,-6) grid (0,0);
      \draw [color=black,->] (-5.0,0) -- (5.0,0);
      \draw [color=black,->] (0,-5.0) -- (0,5.0);

      \filldraw [color={rgb:red,0;green,0;blue,255}] (3,3) circle (2.5pt );
      \filldraw [color={rgb:red,0;green,0;blue,255}] (-1,2) circle (2.5pt );
      \filldraw [color={rgb:red,0;green,0;blue,255}] (-4,-4) circle (2.5pt );
      \filldraw [color={rgb:red,0;green,0;blue,255}] (0,-1) circle (2.5pt );
      \filldraw [color={rgb:red,0;green,0;blue,255}] (4,-3) circle (2.5pt );

      \filldraw [color={rgb:red,255;green,0;blue,0}] (2,1) circle (2.5pt );

      \node at (3.3,2.7) {$x_1$};
      \node at (-0.7,1.7) {$x_2$};
      \node at (-3.7,-4.3) {$x_3$};
      \node at (0.3,-1.3) {$x_4$};
      \node at (4.3,-3.3) {$x_5$};

      \node at (2.3,0.7) {$q$};
      \end{tikzpicture}
  \end{column}
 \end{columns}
\end{frame}

% Beispiel1 Abstand x1
\begin{frame}<handout:0>{Full Search}{Example}
 \begin{columns}[T]
  \begin{column}{.35\textwidth}
    \begin{block}{Result}
      $d(q, x_1) \approx 2.236$
    \end{block}
  \end{column}
  \begin{column}{.6\textwidth}
     \begin{tikzpicture}[%
      scale=0.6,
      ]
      \clip (-5.0,-5.0) rectangle (5.0,5.0);
      \draw [help lines] (0,0) grid (6,5);
      \draw [help lines] (-6,0) grid (0,5);
      \draw [help lines] (0,-6) grid (6,0);
      \draw [help lines] (-6,-6) grid (0,0);
      \draw [color=black,->] (-5.0,0) -- (5.0,0);
      \draw [color=black,->] (0,-5.0) -- (0,5.0);

      \draw[color=gray, line width=0.5pt, dashed] (2,1) -- (3,3);

      \filldraw [color={rgb:red,0;green,0;blue,255}] (3,3) circle (2.5pt );
      \filldraw [color={rgb:red,0;green,0;blue,255}] (-1,2) circle (2.5pt );
      \filldraw [color={rgb:red,0;green,0;blue,255}] (-4,-4) circle (2.5pt );
      \filldraw [color={rgb:red,0;green,0;blue,255}] (0,-1) circle (2.5pt );
      \filldraw [color={rgb:red,0;green,0;blue,255}] (4,-3) circle (2.5pt );

      \filldraw [color={rgb:red,255;green,0;blue,0}] (2,1) circle (2.5pt );

      \node at (3.3,2.7) {$x_1$};
      \node at (-0.7,1.7) {$x_2$};
      \node at (-3.7,-4.3) {$x_3$};
      \node at (0.3,-1.3) {$x_4$};
      \node at (4.3,-3.3) {$x_5$};

      \node at (2.3,0.7) {$q$};
      \end{tikzpicture}
  \end{column}
 \end{columns}
\end{frame}

% Beispiel1 Abstand x2
\begin{frame}<handout:0>{Full Search}{Example}
 \begin{columns}[T]
  \begin{column}{.35\textwidth}
    \begin{block}{Result}
      $d(q, x_1) \approx 2.236$\\
      $d(q, x_2) \approx 3.162$
    \end{block}
  \end{column}
  \begin{column}{.6\textwidth}
     \begin{tikzpicture}[%
      scale=0.6,
      ]
      \clip (-5.0,-5.0) rectangle (5.0,5.0);
      \draw [help lines] (0,0) grid (6,5);
      \draw [help lines] (-6,0) grid (0,5);
      \draw [help lines] (0,-6) grid (6,0);
      \draw [help lines] (-6,-6) grid (0,0);
      \draw [color=black,->] (-5.0,0) -- (5.0,0);
      \draw [color=black,->] (0,-5.0) -- (0,5.0);

      \draw[color=gray, line width=0.5pt, dashed] (2,1) -- (3,3);
      \draw[color=gray, line width=0.5pt, dashed] (2,1) -- (-1,2);

      \filldraw [color={rgb:red,0;green,0;blue,255}] (3,3) circle (2.5pt );
      \filldraw [color={rgb:red,0;green,0;blue,255}] (-1,2) circle (2.5pt );
      \filldraw [color={rgb:red,0;green,0;blue,255}] (-4,-4) circle (2.5pt );
      \filldraw [color={rgb:red,0;green,0;blue,255}] (0,-1) circle (2.5pt );
      \filldraw [color={rgb:red,0;green,0;blue,255}] (4,-3) circle (2.5pt );

      \filldraw [color={rgb:red,255;green,0;blue,0}] (2,1) circle (2.5pt );

      \node at (3.3,2.7) {$x_1$};
      \node at (-0.7,1.7) {$x_2$};
      \node at (-3.7,-4.3) {$x_3$};
      \node at (0.3,-1.3) {$x_4$};
      \node at (4.3,-3.3) {$x_5$};

      \node at (2.3,0.7) {$q$};
      \end{tikzpicture}
  \end{column}
 \end{columns}
\end{frame}

% Beispiel1 Abstand x3
\begin{frame}<handout:0>{Full Search}{Example}
 \begin{columns}[T]
  \begin{column}{.35\textwidth}
    \begin{block}{Result}
      $d(q, x_1) \approx 2.236$\\
      $d(q, x_2) \approx 3.162$\\
      $d(q, x_3) \approx 7.810$
    \end{block}
  \end{column}
  \begin{column}{.6\textwidth}
     \begin{tikzpicture}[%
      scale=0.6,
      ]
      \clip (-5.0,-5.0) rectangle (5.0,5.0);
      \draw [help lines] (0,0) grid (6,5);
      \draw [help lines] (-6,0) grid (0,5);
      \draw [help lines] (0,-6) grid (6,0);
      \draw [help lines] (-6,-6) grid (0,0);
      \draw [color=black,->] (-5.0,0) -- (5.0,0);
      \draw [color=black,->] (0,-5.0) -- (0,5.0);

      \draw[color=gray, line width=0.5pt, dashed] (2,1) -- (3,3);
      \draw[color=gray, line width=0.5pt, dashed] (2,1) -- (-1,2);
      \draw[color=gray, line width=0.5pt, dashed] (2,1) -- (-4,-4);

      \filldraw [color={rgb:red,0;green,0;blue,255}] (3,3) circle (2.5pt );
      \filldraw [color={rgb:red,0;green,0;blue,255}] (-1,2) circle (2.5pt );
      \filldraw [color={rgb:red,0;green,0;blue,255}] (-4,-4) circle (2.5pt );
      \filldraw [color={rgb:red,0;green,0;blue,255}] (0,-1) circle (2.5pt );
      \filldraw [color={rgb:red,0;green,0;blue,255}] (4,-3) circle (2.5pt );

      \filldraw [color={rgb:red,255;green,0;blue,0}] (2,1) circle (2.5pt );

      \node at (3.3,2.7) {$x_1$};
      \node at (-0.7,1.7) {$x_2$};
      \node at (-3.7,-4.3) {$x_3$};
      \node at (0.3,-1.3) {$x_4$};
      \node at (4.3,-3.3) {$x_5$};

      \node at (2.3,0.7) {$q$};
      \end{tikzpicture}
  \end{column}
 \end{columns}
\end{frame}

% Beispiel1 Abstand x4
\begin{frame}<handout:0>{Full Search}{Example}
 \begin{columns}[T]
  \begin{column}{.35\textwidth}
    \begin{block}{Result}
      $d(q, x_1) \approx 2.236$\\
      $d(q, x_2) \approx 3.162$\\
      $d(q, x_3) \approx 7.810$\\
      $d(q, x_4) \approx 2.828$
    \end{block}
  \end{column}
  \begin{column}{.6\textwidth}
     \begin{tikzpicture}[%
      scale=0.6,
      ]
      \clip (-5.0,-5.0) rectangle (5.0,5.0);
      \draw [help lines] (0,0) grid (6,5);
      \draw [help lines] (-6,0) grid (0,5);
      \draw [help lines] (0,-6) grid (6,0);
      \draw [help lines] (-6,-6) grid (0,0);
      \draw [color=black,->] (-5.0,0) -- (5.0,0);
      \draw [color=black,->] (0,-5.0) -- (0,5.0);

      \draw[color=gray, line width=0.5pt, dashed] (2,1) -- (3,3);
      \draw[color=gray, line width=0.5pt, dashed] (2,1) -- (-1,2);
      \draw[color=gray, line width=0.5pt, dashed] (2,1) -- (-4,-4);
      \draw[color=gray, line width=0.5pt, dashed] (2,1) -- (0,-1);

      \filldraw [color={rgb:red,0;green,0;blue,255}] (3,3) circle (2.5pt );
      \filldraw [color={rgb:red,0;green,0;blue,255}] (-1,2) circle (2.5pt );
      \filldraw [color={rgb:red,0;green,0;blue,255}] (-4,-4) circle (2.5pt );
      \filldraw [color={rgb:red,0;green,0;blue,255}] (0,-1) circle (2.5pt );
      \filldraw [color={rgb:red,0;green,0;blue,255}] (4,-3) circle (2.5pt );

      \filldraw [color={rgb:red,255;green,0;blue,0}] (2,1) circle (2.5pt );

      \node at (3.3,2.7) {$x_1$};
      \node at (-0.7,1.7) {$x_2$};
      \node at (-3.7,-4.3) {$x_3$};
      \node at (0.3,-1.3) {$x_4$};
      \node at (4.3,-3.3) {$x_5$};

      \node at (2.3,0.7) {$q$};
      \end{tikzpicture}
  \end{column}
 \end{columns}
\end{frame}

% Beispiel1 Abstand x5
\begin{frame}<handout:0>{Full Search}{Example}
 \begin{columns}[T]
  \begin{column}{.35\textwidth}
    \begin{block}{Result}
      $d(q, x_1) \approx 2.236$\\
      $d(q, x_2) \approx 3.162$\\
      $d(q, x_3) \approx 7.810$\\
      $d(q, x_4) \approx 2.828$\\
      $d(q, x_5) \approx 4.472$
    \end{block}
  \end{column}
  \begin{column}{.6\textwidth}
     \begin{tikzpicture}[%
      scale=0.6,
      ]
      \clip (-5.0,-5.0) rectangle (5.0,5.0);
      \draw [help lines] (0,0) grid (6,5);
      \draw [help lines] (-6,0) grid (0,5);
      \draw [help lines] (0,-6) grid (6,0);
      \draw [help lines] (-6,-6) grid (0,0);
      \draw [color=black,->] (-5.0,0) -- (5.0,0);
      \draw [color=black,->] (0,-5.0) -- (0,5.0);

      \draw[color=gray, line width=0.5pt, dashed] (2,1) -- (3,3);
      \draw[color=gray, line width=0.5pt, dashed] (2,1) -- (-1,2);
      \draw[color=gray, line width=0.5pt, dashed] (2,1) -- (-4,-4);
      \draw[color=gray, line width=0.5pt, dashed] (2,1) -- (0,-1);
      \draw[color=gray, line width=0.5pt, dashed] (2,1) -- (4,-3);

      \filldraw [color={rgb:red,0;green,0;blue,255}] (3,3) circle (2.5pt );
      \filldraw [color={rgb:red,0;green,0;blue,255}] (-1,2) circle (2.5pt );
      \filldraw [color={rgb:red,0;green,0;blue,255}] (-4,-4) circle (2.5pt );
      \filldraw [color={rgb:red,0;green,0;blue,255}] (0,-1) circle (2.5pt );
      \filldraw [color={rgb:red,0;green,0;blue,255}] (4,-3) circle (2.5pt );

      \filldraw [color={rgb:red,255;green,0;blue,0}] (2,1) circle (2.5pt );

      \node at (3.3,2.7) {$x_1$};
      \node at (-0.7,1.7) {$x_2$};
      \node at (-3.7,-4.3) {$x_3$};
      \node at (0.3,-1.3) {$x_4$};
      \node at (4.3,-3.3) {$x_5$};

      \node at (2.3,0.7) {$q$};
      \end{tikzpicture}
  \end{column}
 \end{columns}
\end{frame}

% Beispiel1 Ergebnis
\begin{frame}{Full Search}{Example}
 \begin{columns}[T]
  \begin{column}{.35\textwidth}
    \begin{block}{Result}
      \textcolor{green}{$d(q, x_1) \approx 2.236$}\\
      $d(q, x_2) \approx 3.162$\\
      $d(q, x_3) \approx 7.810$\\
      $d(q, x_4) \approx 2.828$\\
      $d(q, x_5) \approx 4.472$
    \end{block}
  \end{column}
  \begin{column}{.6\textwidth}
     \begin{tikzpicture}[%
      scale=0.6,
      ]
      \clip (-5.0,-5.0) rectangle (5.0,5.0);
      \draw [help lines] (0,0) grid (6,5);
      \draw [help lines] (-6,0) grid (0,5);
      \draw [help lines] (0,-6) grid (6,0);
      \draw [help lines] (-6,-6) grid (0,0);
      \draw [color=black,->] (-5.0,0) -- (5.0,0);
      \draw [color=black,->] (0,-5.0) -- (0,5.0);

      \draw[color=green, line width=0.5pt, solid] (2,1) -- (3,3);
      \draw[color=gray, line width=0.5pt, dashed] (2,1) -- (-1,2);
      \draw[color=gray, line width=0.5pt, dashed] (2,1) -- (-4,-4);
      \draw[color=gray, line width=0.5pt, dashed] (2,1) -- (0,-1);
      \draw[color=gray, line width=0.5pt, dashed] (2,1) -- (4,-3);

      \filldraw [color={rgb:red,0;green,0;blue,255}] (3,3) circle (2.5pt );
      \filldraw [color={rgb:red,0;green,0;blue,255}] (-1,2) circle (2.5pt );
      \filldraw [color={rgb:red,0;green,0;blue,255}] (-4,-4) circle (2.5pt );
      \filldraw [color={rgb:red,0;green,0;blue,255}] (0,-1) circle (2.5pt );
      \filldraw [color={rgb:red,0;green,0;blue,255}] (4,-3) circle (2.5pt );

      \filldraw [color={rgb:red,255;green,0;blue,0}] (2,1) circle (2.5pt );

      \node at (3.3,2.7) {$x_1$};
      \node at (-0.7,1.7) {$x_2$};
      \node at (-3.7,-4.3) {$x_3$};
      \node at (0.3,-1.3) {$x_4$};
      \node at (4.3,-3.3) {$x_5$};

      \node at (2.3,0.7) {$q$};
      \end{tikzpicture}
  \end{column}
 \end{columns}
\end{frame}

% Vor- und Nachteile
\begin{frame}{Full Search}{Advantages and disadvantages}
 \begin{block}{Advantages}
  \begin{itemize}
   \item Easy implementation
   \pause
   \item Works in none metric spaces
  \end{itemize}
 \end{block}
 \pause
 \begin{block}{Disadvantages}
  \begin{itemize}
   \item Large runtime on big data sets and in higher multidimensional spaces
  \end{itemize}
 \end{block}
\end{frame}

\begin{frame}{Metric}
  Given a set $X$. A \textit{Metric} on $X$ is a function\\
  $d:X \times X \to \mathbb{R}$, $(x, y) \mapsto d(x, y)$\\
  with:
 \begin{enumerate}
 \pause
  \item $d(x, y) = 0$ exactly when $x=y$.
  \pause
  \item Symmetry: For all $x, y \in X$ is true $d(x, y) = d(y, x)$.
  \pause
  \item Triangle inequality: For all $x, y, z \in X$ is true $d(x, z) \le d(x, y) + d(y,z)$
 \end{enumerate}
 \flushright{\tiny{[Forster, 2006]}}
\end{frame}

\begin{frame}{Triangle inequality}
  \begin{center}
    \begin{tikzpicture}[%
    scale=1.0,
    ]
    \clip (-4.15325,-0.785024) rectangle (0.889982,2.09932);
    \filldraw [color={rgb:red,0;green,0;blue,255}] (0.387719,-0.137281) circle (2.5pt );
    \filldraw [color={rgb:red,0;green,0;blue,255}] (-3.46333,0.27222) circle (2.5pt );
    \filldraw [color={rgb:red,0;green,0;blue,255}] (-0.413857,1.98864) circle (2.5pt );
    \draw[color={rgb:red,0;green,0;blue,255}, line width=0.5pt, solid] (-3.46333,0.27222) -- (-0.413857,1.98864);
    \draw[color={rgb:red,0;green,0;blue,255}, line width=0.5pt, solid] (-3.46333,0.27222) -- (0.387719,-0.137281);
    \draw[color={rgb:red,0;green,0;blue,255}, line width=0.5pt, solid] (-0.413857,1.98864) -- (0.387719,-0.137281);
    \node at (0.679006,-0.454707) {s};
    \node at (-0.131283,1.69735) {p};
    \node at (-3.18076,-0.163507) {q};
    \end{tikzpicture}
  \end{center}

 \textbf{Lemma} For any $q, s, p \in \mathbb{U}$, $r \in \mathbb{R}$ and $P \subset \mathbb{U}$ is true:
 \begin{enumerate}
 \pause
  \item \label{ineq:1}$|d(p, q) - d(p, s)| \le d(q, s) \le d(p, q) + d(p, s)$
  \pause
  \item \label{ineq:2}$d(q, s) \ge d_P(q, s) := \max_{p \in P}|d(p, q) - d(p, s)|$
  \pause
  \item \label{ineq:3}$d(p, s) > d(p, q) + r \vee d(p, s) < d(p, q) -r \Rightarrow d(q, s) > r$
  \pause
  \item \label{ineq:4}$d(p, s) \ge 2 \cdot d(p, q) \Rightarrow d(q, s) \ge d(q, p)$
 \end{enumerate}
 \flushright{\tiny{[Clarkson, 2005]}}
\end{frame}

\subsection{Orchards Algorithm}
\begin{frame}{Orchards Algorithm}
 \begin{itemize}
  \item Create a list for every item $p \in S$ with all items $x \in S$, ordered ascending to the distance
  \pause
  \item Choose random item $c \in S$ as initial candidate
  \pause
  \item Calculate $d(c, q)$
  \pause
  \item Go along the list of $c$
  \pause
  \item If the current item has smaller distance to $q$ as $c$, choose current item as $c$
  \pause
  \item Abort, if
    \begin{itemize}
      \item at the end of the current list or
      \item $d(c, s) > 2 \cdot d(c, q)$ for the current item of the list (Triangle inequality~\ref{ineq:4})
    \end{itemize}
   \pause
   \item Else $c$ is nearest neighbour
 \end{itemize}
\end{frame}

%Beispiel2
\begin{frame}{Orchards Algorithm}{Example}
 \begin{columns}[T]
  \begin{column}{.35\textwidth}
    $\mathbb{U} = \mathbb{R}^2$
    \begin{block}{Items}
      $x_1 = (3, 3)$\\
      $x_2 = (-1, 2)$\\
      $x_3 = (-4, -4)$\\
      $x_4 = (0, -1)$\\
      $x_5 = (4, -3)$
    \end{block}
    \begin{block}{Query item}
      $q = (2, 1)$
    \end{block}
  \end{column}
  \begin{column}{.60\textwidth}
     \begin{tikzpicture}[%
      scale=0.6,
      ]
      \clip (-5.0,-5.0) rectangle (5.0,5.0);
      \draw [help lines] (0,0) grid (6,5);
      \draw [help lines] (-6,0) grid (0,5);
      \draw [help lines] (0,-6) grid (6,0);
      \draw [help lines] (-6,-6) grid (0,0);
      \draw [color=black,->] (-5.0,0) -- (5.0,0);
      \draw [color=black,->] (0,-5.0) -- (0,5.0);

      \filldraw [color={rgb:red,0;green,0;blue,255}] (3,3) circle (2.5pt );
      \filldraw [color={rgb:red,0;green,0;blue,255}] (-1,2) circle (2.5pt );
      \filldraw [color={rgb:red,0;green,0;blue,255}] (-4,-4) circle (2.5pt );
      \filldraw [color={rgb:red,0;green,0;blue,255}] (0,-1) circle (2.5pt );
      \filldraw [color={rgb:red,0;green,0;blue,255}] (4,-3) circle (2.5pt );

      \filldraw [color={rgb:red,255;green,0;blue,0}] (2,1) circle (2.5pt );

      \node at (3.3,2.7) {$x_1$};
      \node at (-0.7,1.7) {$x_2$};
      \node at (-3.7,-4.3) {$x_3$};
      \node at (0.3,-1.3) {$x_4$};
      \node at (4.3,-3.3) {$x_5$};

      \node at (2.3,0.7) {$q$};
      \end{tikzpicture}
  \end{column}
 \end{columns}
\end{frame}

%Beispiel2 Abstand
\begin{frame}{Orchards Algorithm}{Example}
  \begin{block}{Distances}
  \begin{center}
    \begin{tabular}{r|ccccc}
      & $x_1$ & $x_2$ & $x_3$ & $x_4$ & $x_5$\\
      \hline
      $x_1$ & 0 & $\approx 4.123$ & $\approx 9.899$ & $5$ & $\approx 6.083$\\
      $x_2$ & $\approx 4.123$ & 0 & $\approx 6.708$ & $\approx 3.162$ & $\approx 7.071$\\
      $x_3$ & $\approx 9.899$ & $\approx 6.708$ & 0 & $5$ & $\approx 8.062$\\
      $x_4$ & $5$ & $\approx 3.162$ & $5$ & 0 & $\approx 4.472$\\
      $x_5$ & $\approx 6.083$ & $\approx 7.071$ & $\approx 8.062$ & $\approx 4.472$ & 0\\
    \end{tabular}
  \end{center}
  \end{block}
\end{frame}

%Beispiel2 Listen
\begin{frame}{Orchards Algorithm}{Example}
 \begin{columns}[T]
  \begin{column}{.35\textwidth}
    \begin{block}{Lists}
    $L(x_1) = \{x_2, x_4, x_5, x_3\}$\\
    $L(x_2) = \{x_4, x_1, x_3, x_5\}$\\
    $L(x_3) = \{x_4, x_2, x_5, x_1\}$\\
    $L(x_4) = \{x_2, x_5, x_1, x_3\}$\\
    $L(x_5) = \{x_4, x_1, x_2, x_3\}$\\
    \end{block}
  \end{column}
  \begin{column}{.6\textwidth}
     \begin{tikzpicture}[%
      scale=0.6,
      ]
      \clip (-5.0,-5.0) rectangle (5.0,5.0);
      \draw [help lines] (0,0) grid (6,5);
      \draw [help lines] (-6,0) grid (0,5);
      \draw [help lines] (0,-6) grid (6,0);
      \draw [help lines] (-6,-6) grid (0,0);
      \draw [color=black,->] (-5.0,0) -- (5.0,0);
      \draw [color=black,->] (0,-5.0) -- (0,5.0);

      \draw[color=gray, line width=0.5pt, dashed] (3,3) -- (-1,2);
      \draw[color=gray, line width=0.5pt, dashed] (3,3) -- (-4,-4);
      \draw[color=gray, line width=0.5pt, dashed] (3,3) -- (0,-1);
      \draw[color=gray, line width=0.5pt, dashed] (3,3) -- (4,-3);
      \draw[color=gray, line width=0.5pt, dashed] (-1,2) -- (-4,-4);
      \draw[color=gray, line width=0.5pt, dashed] (-1,2) -- (0,-1);
      \draw[color=gray, line width=0.5pt, dashed] (-1,2) -- (4,-3);
      \draw[color=gray, line width=0.5pt, dashed] (-4,-4) -- (0,-1);
      \draw[color=gray, line width=0.5pt, dashed] (-4,-4) -- (4,-3);
      \draw[color=gray, line width=0.5pt, dashed] (0,-1) -- (4,-3);

      \filldraw [color={rgb:red,0;green,0;blue,255}] (3,3) circle (2.5pt );
      \filldraw [color={rgb:red,0;green,0;blue,255}] (-1,2) circle (2.5pt );
      \filldraw [color={rgb:red,0;green,0;blue,255}] (-4,-4) circle (2.5pt );
      \filldraw [color={rgb:red,0;green,0;blue,255}] (0,-1) circle (2.5pt );
      \filldraw [color={rgb:red,0;green,0;blue,255}] (4,-3) circle (2.5pt );

      \node at (3.3,2.7) {$x_1$};
      \node at (-0.7,1.7) {$x_2$};
      \node at (-3.7,-4.3) {$x_3$};
      \node at (0.3,-1.3) {$x_4$};
      \node at (4.3,-3.3) {$x_5$};
      \end{tikzpicture}
  \end{column}
 \end{columns}
\end{frame}

\begin{frame}<handout:0>{Orchards Algorithm}{Example}
 \begin{columns}[T]
  \begin{column}{.35\textwidth}
    \begin{itemize}
     \item Set $c:=x_3$ and $s:= x_4$
    \end{itemize}
  \end{column}
  \begin{column}{.60\textwidth}
     \begin{tikzpicture}[%
      scale=0.6,
      ]
      \clip (-5.0,-5.0) rectangle (5.0,5.0);
      \draw [help lines] (0,0) grid (6,5);
      \draw [help lines] (-6,0) grid (0,5);
      \draw [help lines] (0,-6) grid (6,0);
      \draw [help lines] (-6,-6) grid (0,0);
      \draw [color=black,->] (-5.0,0) -- (5.0,0);
      \draw [color=black,->] (0,-5.0) -- (0,5.0);

      \draw [color=gray, line width=0.5pt, solid] (-4,-4) circle (5);
      \draw [color=gray, line width=0.5pt, solid] (-4,-4) circle (6.708);
      \draw [color=gray, line width=0.5pt, solid] (-4,-4) circle (8.062);
      \draw [color=gray, line width=0.5pt, solid] (-4,-4) circle (9.899);
      \draw [color=yellow, line width=0.5pt, solid] (-4,-4) circle (15.62);

      \filldraw [color=blue] (3,3) circle (2.5pt );
      \filldraw [color=blue] (-1,2) circle (2.5pt );
      \filldraw [color=blue] (-4,-4) circle (2.5pt );
      \filldraw [color=yellow] (-4,-4) circle (2.1pt );
      \filldraw [color=blue] (0,-1) circle (2.5pt );
      \filldraw [color=white] (0,-1) circle (2.1pt );
      \filldraw [color=blue] (4,-3) circle (2.5pt );

      \filldraw [color=red] (2,1) circle (2.5pt );

      \node at (3.3,2.7) {$x_1$};
      \node at (-0.7,1.7) {$x_2$};
      \node at (-3.7,-4.3) {$x_3$};
      \node at (0.3,-1.3) {$x_4$};
      \node at (4.3,-3.3) {$x_5$};

      \node at (2.3,0.7) {$q$};
      \end{tikzpicture}
  \end{column}
 \end{columns}
\end{frame}

%Beispiel2 Wähle c=x_3 und s=x_4
\begin{frame}{Orchards Algorithm}{Example}
 \begin{columns}[T]
  \begin{column}{.35\textwidth}
    \begin{itemize}
     \item Set $c:=x_3$ and $s:= x_4$
     \item As $7.810 \approx d(c, q) > d(s, q) \approx 2.828$, set $c:=s$
    \end{itemize}
  \end{column}
  \begin{column}{.60\textwidth}
     \begin{tikzpicture}[%
      scale=0.6,
      ]
      \clip (-5.0,-5.0) rectangle (5.0,5.0);
      \draw [help lines] (0,0) grid (6,5);
      \draw [help lines] (-6,0) grid (0,5);
      \draw [help lines] (0,-6) grid (6,0);
      \draw [help lines] (-6,-6) grid (0,0);
      \draw [color=black,->] (-5.0,0) -- (5.0,0);
      \draw [color=black,->] (0,-5.0) -- (0,5.0);

      \draw [color=gray, line width=0.5pt, solid] (-4,-4) circle (5);
      \draw [color=gray, line width=0.5pt, solid] (-4,-4) circle (6.708);
      \draw [color=gray, line width=0.5pt, solid] (-4,-4) circle (8.062);
      \draw [color=gray, line width=0.5pt, solid] (-4,-4) circle (9.899);
      \draw [color=yellow, line width=0.5pt, solid] (-4,-4) circle (15.62);
      \draw [color=yellow, line width=0.5pt, solid] (-4,-4) circle (15.62);

      \draw[color=gray, line width=0.5pt, dashed] (2,1) -- (-4,-4);
      \draw[color=gray, line width=0.5pt, dashed] (2,1) -- (0,-1);

      \filldraw [color=blue] (3,3) circle (2.5pt );
      \filldraw [color=blue] (-1,2) circle (2.5pt );
      \filldraw [color=blue] (-4,-4) circle (2.5pt );
      \filldraw [color=yellow] (-4,-4) circle (2.1pt );
      \filldraw [color=blue] (0,-1) circle (2.5pt );
      \filldraw [color=white] (0,-1) circle (2.1pt );
      \filldraw [color=blue] (4,-3) circle (2.5pt );

      \filldraw [color=red] (2,1) circle (2.5pt );

      \node at (3.3,2.7) {$x_1$};
      \node at (-0.7,1.7) {$x_2$};
      \node at (-3.7,-4.3) {$x_3$};
      \node at (0.3,-1.3) {$x_4$};
      \node at (4.3,-3.3) {$x_5$};

      \node at (2.3,0.7) {$q$};
      \end{tikzpicture}
  \end{column}
 \end{columns}
\end{frame}

%Beispiel2 Wähle c=x_4 und s=x_2
\begin{frame}<handout:0>{Orchards Algorithm}{Example}
 \begin{columns}[T]
  \begin{column}{.35\textwidth}
    \begin{itemize}
     \item Set $c:=x_4$ and $s:= x_2$
    \end{itemize}
  \end{column}
  \begin{column}{.60\textwidth}
     \begin{tikzpicture}[%
      scale=0.6,
      ]
      \clip (-5.0,-5.0) rectangle (5.0,5.0);
      \draw [help lines] (0,0) grid (6,5);
      \draw [help lines] (-6,0) grid (0,5);
      \draw [help lines] (0,-6) grid (6,0);
      \draw [help lines] (-6,-6) grid (0,0);
      \draw [color=black,->] (-5.0,0) -- (5.0,0);
      \draw [color=black,->] (0,-5.0) -- (0,5.0);

      \draw [color=gray, line width=0.5pt, solid] (0,-1) circle (3.162);
      \draw [color=gray, line width=0.5pt, solid] (0,-1) circle (4.472);
      \draw [color=gray, line width=0.5pt, solid] (0,-1) circle (5);
      \draw [color=yellow, line width=0.5pt, solid] (0,-1) circle (5.656);
      \draw [color=yellow, line width=0.5pt, solid] (0,-1) circle (5.656);

      \filldraw [color=blue] (3,3) circle (2.5pt );
      \filldraw [color=blue] (-1,2) circle (2.5pt );
      \filldraw [color=white] (-1,2) circle (2.1pt );
      \filldraw [color=blue] (-4,-4) circle (2.5pt );
      \filldraw [color=blue] (0,-1) circle (2.5pt );
      \filldraw [color=yellow] (0,-1) circle (2.1pt );
      \filldraw [color=blue] (4,-3) circle (2.5pt );

      \filldraw [color=red] (2,1) circle (2.5pt );

      \node at (3.3,2.7) {$x_1$};
      \node at (-0.7,1.7) {$x_2$};
      \node at (-3.7,-4.3) {$x_3$};
      \node at (0.3,-1.3) {$x_4$};
      \node at (4.3,-3.3) {$x_5$};

      \node at (2.3,0.7) {$q$};
      \end{tikzpicture}
  \end{column}
 \end{columns}
\end{frame}

%Beispiel2 Wähle c=x_4 und s=x_2
\begin{frame}<handout:0>{Orchards Algorithm}{Example}
 \begin{columns}[T]
  \begin{column}{.35\textwidth}
    \begin{itemize}
     \item Set $c:=x_4$ and $s:= x_2$
     \item As $2.828 \approx d(c, q) < d(s, q) \approx 3.162$, no new $c$
    \end{itemize}
  \end{column}
  \begin{column}{.60\textwidth}
     \begin{tikzpicture}[%
      scale=0.6,
      ]
      \clip (-5.0,-5.0) rectangle (5.0,5.0);
      \draw [help lines] (0,0) grid (6,5);
      \draw [help lines] (-6,0) grid (0,5);
      \draw [help lines] (0,-6) grid (6,0);
      \draw [help lines] (-6,-6) grid (0,0);
      \draw [color=black,->] (-5.0,0) -- (5.0,0);
      \draw [color=black,->] (0,-5.0) -- (0,5.0);

      \draw [color=gray, line width=0.5pt, solid] (0,-1) circle (3.162);
      \draw [color=gray, line width=0.5pt, solid] (0,-1) circle (4.472);
      \draw [color=gray, line width=0.5pt, solid] (0,-1) circle (5);
      \draw [color=yellow, line width=0.5pt, solid] (0,-1) circle (5.656);

      \draw[color=gray, line width=0.5pt, dashed] (2,1) -- (-1,2);
      \draw[color=gray, line width=0.5pt, dashed] (2,1) -- (0,-1);

      \filldraw [color=blue] (3,3) circle (2.5pt );
      \filldraw [color=blue] (-1,2) circle (2.5pt );
      \filldraw [color=white] (-1,2) circle (2.1pt );
      \filldraw [color=blue] (-4,-4) circle (2.5pt );
      \filldraw [color=blue] (0,-1) circle (2.5pt );
      \filldraw [color=yellow] (0,-1) circle (2.1pt );
      \filldraw [color=blue] (4,-3) circle (2.5pt );

      \filldraw [color=red] (2,1) circle (2.5pt );

      \node at (3.3,2.7) {$x_1$};
      \node at (-0.7,1.7) {$x_2$};
      \node at (-3.7,-4.3) {$x_3$};
      \node at (0.3,-1.3) {$x_4$};
      \node at (4.3,-3.3) {$x_5$};

      \node at (2.3,0.7) {$q$};
      \end{tikzpicture}
  \end{column}
 \end{columns}
\end{frame}

%Beispiel2 Wähle c=x_4 und s=x_2
\begin{frame}{Orchards Algorithm}{Example}
 \begin{columns}[T]
  \begin{column}{.35\textwidth}
    \begin{itemize}
     \item Set $c:=x_4$ and $s:= x_2$
     \item As $2.828 \approx d(c, q) < d(s, q) \approx 3.162$, no new $c$
     \item As $3.162 \approx d(c, s) < 2 \cdot d(c, q) \approx 5.656$, no abort
    \end{itemize}
  \end{column}
  \begin{column}{.60\textwidth}
     \begin{tikzpicture}[%
      scale=0.6,
      ]
      \clip (-5.0,-5.0) rectangle (5.0,5.0);
      \draw [help lines] (0,0) grid (6,5);
      \draw [help lines] (-6,0) grid (0,5);
      \draw [help lines] (0,-6) grid (6,0);
      \draw [help lines] (-6,-6) grid (0,0);
      \draw [color=black,->] (-5.0,0) -- (5.0,0);
      \draw [color=black,->] (0,-5.0) -- (0,5.0);

      \draw [color=gray, line width=0.5pt, solid] (0,-1) circle (3.162);
      \draw [color=gray, line width=0.5pt, solid] (0,-1) circle (4.472);
      \draw [color=gray, line width=0.5pt, solid] (0,-1) circle (5);
      \draw [color=yellow, line width=0.5pt, solid] (0,-1) circle (5.656);

      \draw[color=gray, line width=0.5pt, dashed] (2,1) -- (-1,2);
      \draw[color=gray, line width=0.5pt, dashed] (2,1) -- (0,-1);
      \draw[color=gray, line width=0.5pt, dashed] (-1,2) -- (0,-1);

      \filldraw [color=blue] (3,3) circle (2.5pt );
      \filldraw [color=blue] (-1,2) circle (2.5pt );
      \filldraw [color=white] (-1,2) circle (2.1pt );
      \filldraw [color=blue] (-4,-4) circle (2.5pt );
      \filldraw [color=blue] (0,-1) circle (2.5pt );
      \filldraw [color=yellow] (0,-1) circle (2.1pt );
      \filldraw [color=blue] (4,-3) circle (2.5pt );

      \filldraw [color=red] (2,1) circle (2.5pt );

      \node at (3.3,2.7) {$x_1$};
      \node at (-0.7,1.7) {$x_2$};
      \node at (-3.7,-4.3) {$x_3$};
      \node at (0.3,-1.3) {$x_4$};
      \node at (4.3,-3.3) {$x_5$};

      \node at (2.3,0.7) {$q$};
      \end{tikzpicture}
  \end{column}
 \end{columns}
\end{frame}

%Beispiel2 Wähle c=x_4 und s=x_5
\begin{frame}<handout:0>{Orchards Algorithm}{Example}
 \begin{columns}[T]
  \begin{column}{.35\textwidth}
    \begin{itemize}
     \item Set $s:= x_5$
    \end{itemize}
  \end{column}
  \begin{column}{.60\textwidth}
     \begin{tikzpicture}[%
      scale=0.6,
      ]
      \clip (-5.0,-5.0) rectangle (5.0,5.0);
      \draw [help lines] (0,0) grid (6,5);
      \draw [help lines] (-6,0) grid (0,5);
      \draw [help lines] (0,-6) grid (6,0);
      \draw [help lines] (-6,-6) grid (0,0);
      \draw [color=black,->] (-5.0,0) -- (5.0,0);
      \draw [color=black,->] (0,-5.0) -- (0,5.0);

      \draw [color=gray, line width=0.5pt, solid] (0,-1) circle (3.162);
      \draw [color=gray, line width=0.5pt, solid] (0,-1) circle (4.472);
      \draw [color=gray, line width=0.5pt, solid] (0,-1) circle (5);
      \draw [color=yellow, line width=0.5pt, solid] (0,-1) circle (5.656);

      \filldraw [color=blue] (3,3) circle (2.5pt );
      \filldraw [color=blue] (-1,2) circle (2.5pt );
      \filldraw [color=blue] (-4,-4) circle (2.5pt );
      \filldraw [color=blue] (0,-1) circle (2.5pt );
      \filldraw [color=yellow] (0,-1) circle (2.1pt );
      \filldraw [color=blue] (4,-3) circle (2.5pt );
      \filldraw [color=white] (4,-3) circle (2.1pt );

      \filldraw [color=red] (2,1) circle (2.5pt );

      \node at (3.3,2.7) {$x_1$};
      \node at (-0.7,1.7) {$x_2$};
      \node at (-3.7,-4.3) {$x_3$};
      \node at (0.3,-1.3) {$x_4$};
      \node at (4.3,-3.3) {$x_5$};

      \node at (2.3,0.7) {$q$};
      \end{tikzpicture}
  \end{column}
 \end{columns}
\end{frame}

%Beispiel2 Wähle c=x_4 und s=x_5
\begin{frame}<handout:0>{Orchards Algorithm}{Example}
 \begin{columns}[T]
  \begin{column}{.35\textwidth}
    \begin{itemize}
     \item Set $s:= x_5$
     \item As $2.828 \approx d(c, q) < d(s, q) \approx 4.472$, no new $c$
    \end{itemize}
  \end{column}
  \begin{column}{.60\textwidth}
     \begin{tikzpicture}[%
      scale=0.6,
      ]
      \clip (-5.0,-5.0) rectangle (5.0,5.0);
      \draw [help lines] (0,0) grid (6,5);
      \draw [help lines] (-6,0) grid (0,5);
      \draw [help lines] (0,-6) grid (6,0);
      \draw [help lines] (-6,-6) grid (0,0);
      \draw [color=black,->] (-5.0,0) -- (5.0,0);
      \draw [color=black,->] (0,-5.0) -- (0,5.0);

      \draw [color=gray, line width=0.5pt, solid] (0,-1) circle (3.162);
      \draw [color=gray, line width=0.5pt, solid] (0,-1) circle (4.472);
      \draw [color=gray, line width=0.5pt, solid] (0,-1) circle (5);
      \draw [color=yellow, line width=0.5pt, solid] (0,-1) circle (5.656);

      \draw[color=gray, line width=0.5pt, dashed] (2,1) -- (4,-3);
      \draw[color=gray, line width=0.5pt, dashed] (2,1) -- (0,-1);

      \filldraw [color=blue] (3,3) circle (2.5pt );
      \filldraw [color=blue] (-1,2) circle (2.5pt );
      \filldraw [color=blue] (-4,-4) circle (2.5pt );
      \filldraw [color=blue] (0,-1) circle (2.5pt );
      \filldraw [color=yellow] (0,-1) circle (2.1pt );
      \filldraw [color=blue] (4,-3) circle (2.5pt );
      \filldraw [color=white] (4,-3) circle (2.1pt );

      \filldraw [color=red] (2,1) circle (2.5pt );

      \node at (3.3,2.7) {$x_1$};
      \node at (-0.7,1.7) {$x_2$};
      \node at (-3.7,-4.3) {$x_3$};
      \node at (0.3,-1.3) {$x_4$};
      \node at (4.3,-3.3) {$x_5$};

      \node at (2.3,0.7) {$q$};
      \end{tikzpicture}
  \end{column}
 \end{columns}
\end{frame}

%Beispiel2 Wähle c=x_4 und s=x_5
\begin{frame}{Orchards Algorithm}{Example}
 \begin{columns}[T]
  \begin{column}{.35\textwidth}
    \begin{itemize}
     \item Set $s:= x_5$
     \item As $2.828 \approx d(c, q) < d(s, q) \approx 4.472$, no new $c$
     \item As $4.472 \approx d(c, s) < 2 \cdot d(c, q) \approx 5.656$, no abort
    \end{itemize}
  \end{column}
  \begin{column}{.60\textwidth}
     \begin{tikzpicture}[%
      scale=0.6,
      ]
      \clip (-5.0,-5.0) rectangle (5.0,5.0);
      \draw [help lines] (0,0) grid (6,5);
      \draw [help lines] (-6,0) grid (0,5);
      \draw [help lines] (0,-6) grid (6,0);
      \draw [help lines] (-6,-6) grid (0,0);
      \draw [color=black,->] (-5.0,0) -- (5.0,0);
      \draw [color=black,->] (0,-5.0) -- (0,5.0);

      \draw [color=gray, line width=0.5pt, solid] (0,-1) circle (3.162);
      \draw [color=gray, line width=0.5pt, solid] (0,-1) circle (4.472);
      \draw [color=gray, line width=0.5pt, solid] (0,-1) circle (5);
      \draw [color=yellow, line width=0.5pt, solid] (0,-1) circle (5.656);

      \draw[color=gray, line width=0.5pt, dashed] (2,1) -- (4,-3);
      \draw[color=gray, line width=0.5pt, dashed] (2,1) -- (0,-1);
      \draw[color=gray, line width=0.5pt, dashed] (4,-3) -- (0,-1);

      \filldraw [color=blue] (3,3) circle (2.5pt );
      \filldraw [color=blue] (-1,2) circle (2.5pt );
      \filldraw [color=blue] (-4,-4) circle (2.5pt );
      \filldraw [color=blue] (0,-1) circle (2.5pt );
      \filldraw [color=yellow] (0,-1) circle (2.1pt );
      \filldraw [color=blue] (4,-3) circle (2.5pt );
      \filldraw [color=white] (4,-3) circle (2.1pt );

      \filldraw [color=red] (2,1) circle (2.5pt );

      \node at (3.3,2.7) {$x_1$};
      \node at (-0.7,1.7) {$x_2$};
      \node at (-3.7,-4.3) {$x_3$};
      \node at (0.3,-1.3) {$x_4$};
      \node at (4.3,-3.3) {$x_5$};

      \node at (2.3,0.7) {$q$};
      \end{tikzpicture}
  \end{column}
 \end{columns}
\end{frame}

%Beispiel2 Wähle c=x_4 und s=x_1
\begin{frame}<handout:0>{Orchards Algorithm}{Example}
 \begin{columns}[T]
  \begin{column}{.35\textwidth}
    \begin{itemize}
     \item Set $s:= x_1$
    \end{itemize}
  \end{column}
  \begin{column}{.60\textwidth}
     \begin{tikzpicture}[%
      scale=0.6,
      ]
      \clip (-5.0,-5.0) rectangle (5.0,5.0);
      \draw [help lines] (0,0) grid (6,5);
      \draw [help lines] (-6,0) grid (0,5);
      \draw [help lines] (0,-6) grid (6,0);
      \draw [help lines] (-6,-6) grid (0,0);
      \draw [color=black,->] (-5.0,0) -- (5.0,0);
      \draw [color=black,->] (0,-5.0) -- (0,5.0);

      \draw [color=gray, line width=0.5pt, solid] (0,-1) circle (3.162);
      \draw [color=gray, line width=0.5pt, solid] (0,-1) circle (4.472);
      \draw [color=gray, line width=0.5pt, solid] (0,-1) circle (5);

      \filldraw [color=blue] (3,3) circle (2.5pt );
      \filldraw [color=white] (3,3) circle (2.1pt );
      \filldraw [color=blue] (-1,2) circle (2.5pt );
      \filldraw [color=blue] (-4,-4) circle (2.5pt );
      \filldraw [color=blue] (0,-1) circle (2.5pt );
      \filldraw [color=yellow] (0,-1) circle (2.1pt );
      \filldraw [color=blue] (4,-3) circle (2.5pt );

      \filldraw [color=red] (2,1) circle (2.5pt );

      \node at (3.3,2.7) {$x_1$};
      \node at (-0.7,1.7) {$x_2$};
      \node at (-3.7,-4.3) {$x_3$};
      \node at (0.3,-1.3) {$x_4$};
      \node at (4.3,-3.3) {$x_5$};

      \node at (2.3,0.7) {$q$};
      \end{tikzpicture}
  \end{column}
 \end{columns}
\end{frame}

%Beispiel2 Wähle c=x_4 und s=x_1
\begin{frame}{Orchards Algorithm}{Example}
 \begin{columns}[T]
  \begin{column}{.35\textwidth}
    \begin{itemize}
     \item Set $s:= x_1$
     \item As $2.828 \approx d(c, q) > d(s, q) \approx 2.236$, set $c:=s$
    \end{itemize}
  \end{column}
  \begin{column}{.60\textwidth}
     \begin{tikzpicture}[%
      scale=0.6,
      ]
      \clip (-5.0,-5.0) rectangle (5.0,5.0);
      \draw [help lines] (0,0) grid (6,5);
      \draw [help lines] (-6,0) grid (0,5);
      \draw [help lines] (0,-6) grid (6,0);
      \draw [help lines] (-6,-6) grid (0,0);
      \draw [color=black,->] (-5.0,0) -- (5.0,0);
      \draw [color=black,->] (0,-5.0) -- (0,5.0);

      \draw [color=gray, line width=0.5pt, solid] (0,-1) circle (3.162);
      \draw [color=gray, line width=0.5pt, solid] (0,-1) circle (4.472);
      \draw [color=gray, line width=0.5pt, solid] (0,-1) circle (5);

      \draw[color=gray, line width=0.5pt, dashed] (2,1) -- (3,3);
      \draw[color=gray, line width=0.5pt, dashed] (2,1) -- (0,-1);

      \filldraw [color=blue] (3,3) circle (2.5pt );
      \filldraw [color=white] (3,3) circle (2.1pt );
      \filldraw [color=blue] (-1,2) circle (2.5pt );
      \filldraw [color=blue] (-4,-4) circle (2.5pt );
      \filldraw [color=blue] (0,-1) circle (2.5pt );
      \filldraw [color=yellow] (0,-1) circle (2.1pt );
      \filldraw [color=blue] (4,-3) circle (2.5pt );

      \filldraw [color=red] (2,1) circle (2.5pt );

      \node at (3.3,2.7) {$x_1$};
      \node at (-0.7,1.7) {$x_2$};
      \node at (-3.7,-4.3) {$x_3$};
      \node at (0.3,-1.3) {$x_4$};
      \node at (4.3,-3.3) {$x_5$};

      \node at (2.3,0.7) {$q$};
      \end{tikzpicture}
  \end{column}
 \end{columns}
\end{frame}

%Beispiel2 Wähle c=x_1 und s=x_2
\begin{frame}<handout:0>{Orchards Algorithm}{Example}
 \begin{columns}[T]
  \begin{column}{.35\textwidth}
    \begin{itemize}
     \item Set $c:= x_1$ and $s:=x_2$
    \end{itemize}
  \end{column}
  \begin{column}{.60\textwidth}
     \begin{tikzpicture}[%
      scale=0.6,
      ]
      \clip (-5.0,-5.0) rectangle (5.0,5.0);
      \draw [help lines] (0,0) grid (6,5);
      \draw [help lines] (-6,0) grid (0,5);
      \draw [help lines] (0,-6) grid (6,0);
      \draw [help lines] (-6,-6) grid (0,0);
      \draw [color=black,->] (-5.0,0) -- (5.0,0);
      \draw [color=black,->] (0,-5.0) -- (0,5.0);

      \draw [color=gray, line width=0.5pt, solid] (3,3) circle (4.123);
      \draw [color=gray, line width=0.5pt, solid] (3,3) circle (5);
      \draw [color=gray, line width=0.5pt, solid] (3,3) circle (6.083);
      \draw [color=gray, line width=0.5pt, solid] (3,3) circle (9.899);
      \draw [color=yellow, line width=0.5pt, solid] (3,3) circle (4.472);

      \filldraw [color=blue] (3,3) circle (2.5pt );
      \filldraw [color=yellow] (3,3) circle (2.1pt );
      \filldraw [color=blue] (-1,2) circle (2.5pt );
      \filldraw [color=white] (-1,2) circle (2.1pt );
      \filldraw [color=blue] (-4,-4) circle (2.5pt );
      \filldraw [color=blue] (0,-1) circle (2.5pt );
      \filldraw [color=blue] (4,-3) circle (2.5pt );

      \filldraw [color=red] (2,1) circle (2.5pt );

      \node at (3.3,2.7) {$x_1$};
      \node at (-0.7,1.7) {$x_2$};
      \node at (-3.7,-4.3) {$x_3$};
      \node at (0.3,-1.3) {$x_4$};
      \node at (4.3,-3.3) {$x_5$};

      \node at (2.3,0.7) {$q$};
      \end{tikzpicture}
  \end{column}
 \end{columns}
\end{frame}

%Beispiel2 Wähle c=x_1 und s=x_2
\begin{frame}<handout:0>{Orchards Algorithm}{Example}
 \begin{columns}[T]
  \begin{column}{.35\textwidth}
    \begin{itemize}
     \item Set $c:= x_1$ and $s:=x_2$
     \item As $2.236 \approx d(c, q) < d(s, q) \approx 3.162$, no new $c$
    \end{itemize}
  \end{column}
  \begin{column}{.60\textwidth}
     \begin{tikzpicture}[%
      scale=0.6,
      ]
      \clip (-5.0,-5.0) rectangle (5.0,5.0);
      \draw [help lines] (0,0) grid (6,5);
      \draw [help lines] (-6,0) grid (0,5);
      \draw [help lines] (0,-6) grid (6,0);
      \draw [help lines] (-6,-6) grid (0,0);
      \draw [color=black,->] (-5.0,0) -- (5.0,0);
      \draw [color=black,->] (0,-5.0) -- (0,5.0);

      \draw [color=gray, line width=0.5pt, solid] (3,3) circle (4.123);
      \draw [color=gray, line width=0.5pt, solid] (3,3) circle (5);
      \draw [color=gray, line width=0.5pt, solid] (3,3) circle (6.083);
      \draw [color=gray, line width=0.5pt, solid] (3,3) circle (9.899);
      \draw [color=yellow, line width=0.5pt, solid] (3,3) circle (4.472);

      \draw[color=gray, line width=0.5pt, dashed] (2,1) -- (3,3);
      \draw[color=gray, line width=0.5pt, dashed] (2,1) -- (-1,2);

      \filldraw [color=blue] (3,3) circle (2.5pt );
      \filldraw [color=yellow] (3,3) circle (2.1pt );
      \filldraw [color=blue] (-1,2) circle (2.5pt );
      \filldraw [color=white] (-1,2) circle (2.1pt );
      \filldraw [color=blue] (-4,-4) circle (2.5pt );
      \filldraw [color=blue] (0,-1) circle (2.5pt );
      \filldraw [color=blue] (4,-3) circle (2.5pt );

      \filldraw [color=red] (2,1) circle (2.5pt );

      \node at (3.3,2.7) {$x_1$};
      \node at (-0.7,1.7) {$x_2$};
      \node at (-3.7,-4.3) {$x_3$};
      \node at (0.3,-1.3) {$x_4$};
      \node at (4.3,-3.3) {$x_5$};

      \node at (2.3,0.7) {$q$};
      \end{tikzpicture}
  \end{column}
 \end{columns}
\end{frame}

%Beispiel2 Wähle c=x_1 und s=x_2
\begin{frame}{Orchards Algorithm}{Example}
 \begin{columns}[T]
  \begin{column}{.35\textwidth}
    \begin{itemize}
     \item Set $c:= x_1$ and $s:=x_2$
     \item As $2.236 \approx d(c, q) < d(s, q) \approx 3.162$, no new $c$
     \item As $4.123 \approx d(c, s) < 2 \cdot d(c, q) \approx 4.472$, no abort
    \end{itemize}
  \end{column}
  \begin{column}{.60\textwidth}
     \begin{tikzpicture}[%
      scale=0.6,
      ]
      \clip (-5.0,-5.0) rectangle (5.0,5.0);
      \draw [help lines] (0,0) grid (6,5);
      \draw [help lines] (-6,0) grid (0,5);
      \draw [help lines] (0,-6) grid (6,0);
      \draw [help lines] (-6,-6) grid (0,0);
      \draw [color=black,->] (-5.0,0) -- (5.0,0);
      \draw [color=black,->] (0,-5.0) -- (0,5.0);

      \draw [color=gray, line width=0.5pt, solid] (3,3) circle (4.123);
      \draw [color=gray, line width=0.5pt, solid] (3,3) circle (5);
      \draw [color=gray, line width=0.5pt, solid] (3,3) circle (6.083);
      \draw [color=gray, line width=0.5pt, solid] (3,3) circle (9.899);
      \draw [color=yellow, line width=0.5pt, solid] (3,3) circle (4.472);

      \draw[color=gray, line width=0.5pt, dashed] (2,1) -- (3,3);
      \draw[color=gray, line width=0.5pt, dashed] (2,1) -- (-1,2);
      \draw[color=gray, line width=0.5pt, dashed] (3,3) -- (-1,2);

      \filldraw [color=blue] (3,3) circle (2.5pt );
      \filldraw [color=yellow] (3,3) circle (2.1pt );
      \filldraw [color=blue] (-1,2) circle (2.5pt );
      \filldraw [color=white] (-1,2) circle (2.1pt );
      \filldraw [color=blue] (-4,-4) circle (2.5pt );
      \filldraw [color=blue] (0,-1) circle (2.5pt );
      \filldraw [color=blue] (4,-3) circle (2.5pt );

      \filldraw [color=red] (2,1) circle (2.5pt );

      \node at (3.3,2.7) {$x_1$};
      \node at (-0.7,1.7) {$x_2$};
      \node at (-3.7,-4.3) {$x_3$};
      \node at (0.3,-1.3) {$x_4$};
      \node at (4.3,-3.3) {$x_5$};

      \node at (2.3,0.7) {$q$};
      \end{tikzpicture}
  \end{column}
 \end{columns}
\end{frame}

%Beispiel2 Wähle c=x_1 und s=x_4
\begin{frame}<handout:0>{Orchards Algorithm}{Example}
 \begin{columns}[T]
  \begin{column}{.35\textwidth}
    \begin{itemize}
     \item Set $s:=x_4$
    \end{itemize}
  \end{column}
  \begin{column}{.60\textwidth}
     \begin{tikzpicture}[%
      scale=0.6,
      ]
      \clip (-5.0,-5.0) rectangle (5.0,5.0);
      \draw [help lines] (0,0) grid (6,5);
      \draw [help lines] (-6,0) grid (0,5);
      \draw [help lines] (0,-6) grid (6,0);
      \draw [help lines] (-6,-6) grid (0,0);
      \draw [color=black,->] (-5.0,0) -- (5.0,0);
      \draw [color=black,->] (0,-5.0) -- (0,5.0);

      \draw [color=gray, line width=0.5pt, solid] (3,3) circle (4.123);
      \draw [color=gray, line width=0.5pt, solid] (3,3) circle (5);
      \draw [color=gray, line width=0.5pt, solid] (3,3) circle (6.083);
      \draw [color=gray, line width=0.5pt, solid] (3,3) circle (9.899);
      \draw [color=yellow, line width=0.5pt, solid] (3,3) circle (4.472);

      \filldraw [color=blue] (3,3) circle (2.5pt );
      \filldraw [color=yellow] (3,3) circle (2.1pt );
      \filldraw [color=blue] (-1,2) circle (2.5pt );
      \filldraw [color=blue] (-4,-4) circle (2.5pt );
      \filldraw [color=blue] (0,-1) circle (2.5pt );
      \filldraw [color=white] (0,-1) circle (2.1pt );
      \filldraw [color=blue] (4,-3) circle (2.5pt );

      \filldraw [color=red] (2,1) circle (2.5pt );

      \node at (3.3,2.7) {$x_1$};
      \node at (-0.7,1.7) {$x_2$};
      \node at (-3.7,-4.3) {$x_3$};
      \node at (0.3,-1.3) {$x_4$};
      \node at (4.3,-3.3) {$x_5$};

      \node at (2.3,0.7) {$q$};
      \end{tikzpicture}
  \end{column}
 \end{columns}
\end{frame}

%Beispiel2 Wähle c=x_1 und s=x_4
\begin{frame}<handout:0>{Orchards Algorithm}{Example}
 \begin{columns}[T]
  \begin{column}{.35\textwidth}
    \begin{itemize}
     \item Set $s:=x_4$
     \item As $2.236 \approx d(c, q) < d(s, q) \approx 2.828$, no new $c$
    \end{itemize}
  \end{column}
  \begin{column}{.60\textwidth}
     \begin{tikzpicture}[%
      scale=0.6,
      ]
      \clip (-5.0,-5.0) rectangle (5.0,5.0);
      \draw [help lines] (0,0) grid (6,5);
      \draw [help lines] (-6,0) grid (0,5);
      \draw [help lines] (0,-6) grid (6,0);
      \draw [help lines] (-6,-6) grid (0,0);
      \draw [color=black,->] (-5.0,0) -- (5.0,0);
      \draw [color=black,->] (0,-5.0) -- (0,5.0);

      \draw [color=gray, line width=0.5pt, solid] (3,3) circle (4.123);
      \draw [color=gray, line width=0.5pt, solid] (3,3) circle (5);
      \draw [color=gray, line width=0.5pt, solid] (3,3) circle (6.083);
      \draw [color=gray, line width=0.5pt, solid] (3,3) circle (9.899);
      \draw [color=yellow, line width=0.5pt, solid] (3,3) circle (4.472);

      \draw[color=gray, line width=0.5pt, dashed] (2,1) -- (3,3);
      \draw[color=gray, line width=0.5pt, dashed] (2,1) -- (0,-1);

      \filldraw [color=blue] (3,3) circle (2.5pt );
      \filldraw [color=yellow] (3,3) circle (2.1pt );
      \filldraw [color=blue] (-1,2) circle (2.5pt );
      \filldraw [color=blue] (-4,-4) circle (2.5pt );
      \filldraw [color=blue] (0,-1) circle (2.5pt );
      \filldraw [color=white] (0,-1) circle (2.1pt );
      \filldraw [color=blue] (4,-3) circle (2.5pt );

      \filldraw [color=red] (2,1) circle (2.5pt );

      \node at (3.3,2.7) {$x_1$};
      \node at (-0.7,1.7) {$x_2$};
      \node at (-3.7,-4.3) {$x_3$};
      \node at (0.3,-1.3) {$x_4$};
      \node at (4.3,-3.3) {$x_5$};

      \node at (2.3,0.7) {$q$};
      \end{tikzpicture}
  \end{column}
 \end{columns}
\end{frame}

%Beispiel2 Wähle c=x_1 und s=x_4
\begin{frame}{Orchards Algorithm}{Example}
 \begin{columns}[T]
  \begin{column}{.35\textwidth}
    \begin{itemize}
     \item Set $s:=x_4$
     \item As $2.236 \approx d(c, q) < d(s, q) \approx 2.828$, no new $c$
     \item As $5 \approx d(c, s) > 2 \cdot d(c, q) \approx 4.472$, \textcolor{green}{abort}
    \end{itemize}
  \end{column}
  \begin{column}{.60\textwidth}
     \begin{tikzpicture}[%
      scale=0.6,
      ]
      \clip (-5.0,-5.0) rectangle (5.0,5.0);
      \draw [help lines] (0,0) grid (6,5);
      \draw [help lines] (-6,0) grid (0,5);
      \draw [help lines] (0,-6) grid (6,0);
      \draw [help lines] (-6,-6) grid (0,0);
      \draw [color=black,->] (-5.0,0) -- (5.0,0);
      \draw [color=black,->] (0,-5.0) -- (0,5.0);

      \draw [color=gray, line width=0.5pt, solid] (3,3) circle (4.123);
      \draw [color=gray, line width=0.5pt, solid] (3,3) circle (5);
      \draw [color=gray, line width=0.5pt, solid] (3,3) circle (6.083);
      \draw [color=gray, line width=0.5pt, solid] (3,3) circle (9.899);
      \draw [color=yellow, line width=0.5pt, solid] (3,3) circle (4.472);

      \draw[color=gray, line width=0.5pt, dashed] (2,1) -- (3,3);
      \draw[color=gray, line width=0.5pt, dashed] (2,1) -- (0,-1);
      \draw[color=gray, line width=0.5pt, dashed] (3,3) -- (0,-1);

      \filldraw [color=blue] (3,3) circle (2.5pt );
      \filldraw [color=yellow] (3,3) circle (2.1pt );
      \filldraw [color=blue] (-1,2) circle (2.5pt );
      \filldraw [color=blue] (-4,-4) circle (2.5pt );
      \filldraw [color=blue] (0,-1) circle (2.5pt );
      \filldraw [color=white] (0,-1) circle (2.1pt );
      \filldraw [color=blue] (4,-3) circle (2.5pt );

      \filldraw [color=red] (2,1) circle (2.5pt );

      \node at (3.3,2.7) {$x_1$};
      \node at (-0.7,1.7) {$x_2$};
      \node at (-3.7,-4.3) {$x_3$};
      \node at (0.3,-1.3) {$x_4$};
      \node at (4.3,-3.3) {$x_5$};

      \node at (2.3,0.7) {$q$};
      \end{tikzpicture}
  \end{column}
 \end{columns}
\end{frame}

% Vor- und Nachteile
\begin{frame}{Orchards Algorithm}{Advantages and disadvantages}
 \begin{block}{Advantages}
  \begin{itemize}
   \item Faster as Full Search
  \end{itemize}
 \end{block}
 \pause
 \begin{block}{Disadvantages}
  \begin{itemize}
   \item Preprocessing needs large memory and runtime
  \end{itemize}
 \end{block}
 \pause
 \begin{block}{Improvement}
  \begin{itemize}
   \item Use MarkBits to ensure that no distance is calculated twice
  \end{itemize}
 \end{block}
\end{frame}

\subsection{Annulus Method}
\begin{frame}{Annulus Method}
 \begin{itemize}
  \item Create a list for a random item $p^* \in S$ with all items $x \in S$, ordered ascending to the distance
  \pause
  \item Choose random item $c \in S$
  \pause
  \item Walk alternating away from $p^*$ and back to it in the list
  \pause
  \item If current item $s$ has smaller distance to $q$ as $c$, set $c:= s$
  \pause
  \item Current item $s$ is under $c$:
  \begin{itemize}
   \item If $d(p^*, s) < d(p^*, q) - d(c, q)$, ignore all items under $s$ (Triangle inequality~\ref{ineq:3})
  \end{itemize}
  \pause
  \item Current item $s$ is above $c$:
  \begin{itemize}
   \item If $d(p^*, s) > d(p*, q) + d(c, q)$, ignore all items above $s$ (Triangle inequality~\ref{ineq:3})
  \end{itemize}
  \pause
  \item $c$ is the nearest neighbour if the entire list is traversed
 \end{itemize}
\end{frame}

%Beispiel3
\begin{frame}{Annulus Method}{Example}
 \begin{columns}[T]
  \begin{column}{.35\textwidth}
    $\mathbb{U} = \mathbb{R}^2$
    \begin{block}{Items}
      $x_1 = (3, 3)$\\
      $x_2 = (-1, 2)$\\
      $x_3 = (-4, -4)$\\
      $x_4 = (0, -1)$\\
      $x_5 = (4, -3)$
    \end{block}
    \begin{block}{Query item}
      $q = (2, 1)$
    \end{block}
  \end{column}
  \begin{column}{.60\textwidth}
     \begin{tikzpicture}[%
      scale=0.6,
      ]
      \clip (-5.0,-5.0) rectangle (5.0,5.0);
      \draw [help lines] (0,0) grid (6,5);
      \draw [help lines] (-6,0) grid (0,5);
      \draw [help lines] (0,-6) grid (6,0);
      \draw [help lines] (-6,-6) grid (0,0);
      \draw [color=black,->] (-5.0,0) -- (5.0,0);
      \draw [color=black,->] (0,-5.0) -- (0,5.0);

      \filldraw [color={rgb:red,0;green,0;blue,255}] (3,3) circle (2.5pt );
      \filldraw [color={rgb:red,0;green,0;blue,255}] (-1,2) circle (2.5pt );
      \filldraw [color={rgb:red,0;green,0;blue,255}] (-4,-4) circle (2.5pt );
      \filldraw [color={rgb:red,0;green,0;blue,255}] (0,-1) circle (2.5pt );
      \filldraw [color={rgb:red,0;green,0;blue,255}] (4,-3) circle (2.5pt );

      \filldraw [color={rgb:red,255;green,0;blue,0}] (2,1) circle (2.5pt );

      \node at (3.3,2.7) {$x_1$};
      \node at (-0.7,1.7) {$x_2$};
      \node at (-3.7,-4.3) {$x_3$};
      \node at (0.3,-1.3) {$x_4$};
      \node at (4.3,-3.3) {$x_5$};

      \node at (2.3,0.7) {$q$};
      \end{tikzpicture}
  \end{column}
 \end{columns}
\end{frame}

%Beispiel3 Abstand
\begin{frame}{Annulus Method}{Example}
  \begin{block}{Distances}
  \begin{center}
    \begin{tabular}{r|ccccc}
      & $x_1$ & $x_2$ & $x_3$ & $x_4$ & $x_5$\\
      \hline
      $x_5$ & $\approx 6.083$ & $\approx 7.071$ & $\approx 8.062$ & $\approx 4.472$ & 0\\
    \end{tabular}
  \end{center}
  \end{block}
\end{frame}

%Beispiel3 Listen
\begin{frame}{Annulus Method}{Example}
 \begin{columns}[T]
  \begin{column}{.35\textwidth}
    $p^*:=x_5$ with $d(p^*, q) \approx 4.472$
    \begin{block}{List}
    $L(x_5) = \{x_5, x_4, x_1, x_2, x_3\}$\\
    \end{block}
  \end{column}
  \begin{column}{.6\textwidth}
     \begin{tikzpicture}[%
      scale=0.6,
      ]
      \clip (-5.0,-5.0) rectangle (5.0,5.0);
      \draw [help lines] (0,0) grid (6,5);
      \draw [help lines] (-6,0) grid (0,5);
      \draw [help lines] (0,-6) grid (6,0);
      \draw [help lines] (-6,-6) grid (0,0);
      \draw [color=black,->] (-5.0,0) -- (5.0,0);
      \draw [color=black,->] (0,-5.0) -- (0,5.0);

      \draw[color=gray, line width=0.5pt, dashed] (3,3) -- (4,-3);
      \draw[color=gray, line width=0.5pt, dashed] (4,-3) -- (-4,-4);
      \draw[color=gray, line width=0.5pt, dashed] (4,-3) -- (0,-1);
      \draw[color=gray, line width=0.5pt, dashed] (4,-3) -- (-1,2);

      \filldraw [color=blue] (3,3) circle (2.5pt );
      \filldraw [color=blue] (-1,2) circle (2.5pt );
      \filldraw [color=blue] (-4,-4) circle (2.5pt );
      \filldraw [color=blue] (0,-1) circle (2.5pt );
      \filldraw [color=blue] (4,-3) circle (2.5pt );
      \filldraw [color=yellow] (4,-3) circle (2.1pt );

      \node at (3.3,2.7) {$x_1$};
      \node at (-0.7,1.7) {$x_2$};
      \node at (-3.7,-4.3) {$x_3$};
      \node at (0.3,-1.3) {$x_4$};
      \node at (4.3,-3.3) {$x_5$};
      \end{tikzpicture}
  \end{column}
 \end{columns}
\end{frame}

%Beispiel3 c:=x_2
\begin{frame}<handout:0>{Annulus Method}{Example}
 \begin{columns}[T]
  \begin{column}{.35\textwidth}
    \begin{itemize}
     \item Set $c:=x_2$ with $d(c,q) \approx 3.162$
    \end{itemize}
  \end{column}
  \begin{column}{.60\textwidth}
     \begin{tikzpicture}[%
      scale=0.6,
      ]
      \clip (-5.0,-5.0) rectangle (5.0,5.0);
      \draw [help lines] (0,0) grid (6,5);
      \draw [help lines] (-6,0) grid (0,5);
      \draw [help lines] (0,-6) grid (6,0);
      \draw [help lines] (-6,-6) grid (0,0);
      \draw [color=black,->] (-5.0,0) -- (5.0,0);
      \draw [color=black,->] (0,-5.0) -- (0,5.0);

      \draw [color=gray, line width=0.5pt, solid] (4,-3) circle (6.083);
      \draw [color=gray, line width=0.5pt, solid] (4,-3) circle (7.071);
      \draw [color=gray, line width=0.5pt, solid] (4,-3) circle (8.062);
      \draw [color=gray, line width=0.5pt, solid] (4,-3) circle (4.472);

      \filldraw [color=blue] (3,3) circle (2.5pt );
      \filldraw [color=blue] (-1,2) circle (2.5pt );
      \filldraw [color=white] (-1,2) circle (2.1pt );
      \filldraw [color=blue] (-4,-4) circle (2.5pt );
      \filldraw [color=blue] (0,-1) circle (2.5pt );
      \filldraw [color=blue] (4,-3) circle (2.5pt );
      \filldraw [color=yellow] (4,-3) circle (2.1pt );

      \filldraw [color=red] (2,1) circle (2.5pt );

      \node at (3.3,2.7) {$x_1$};
      \node at (-0.7,1.7) {$x_2$};
      \node at (-3.7,-4.3) {$x_3$};
      \node at (0.3,-1.3) {$x_4$};
      \node at (4.3,-3.3) {$x_5$};

      \node at (2.3,0.7) {$q$};
      \end{tikzpicture}
  \end{column}
 \end{columns}
\end{frame}

%Beispiel3 c:=x_2
\begin{frame}<handout:0>{Annulus Method}{Example}
 \begin{columns}[T]
  \begin{column}{.35\textwidth}
    \begin{itemize}
     \item Set $c:=x_2$ with $d(c, q) \approx 3.162$
     \item Set $s:=x_3$ with $d(s, q) \approx 7.810$
    \end{itemize}
  \end{column}
  \begin{column}{.60\textwidth}
     \begin{tikzpicture}[%
      scale=0.6,
      ]
      \clip (-5.0,-5.0) rectangle (5.0,5.0);
      \draw [help lines] (0,0) grid (6,5);
      \draw [help lines] (-6,0) grid (0,5);
      \draw [help lines] (0,-6) grid (6,0);
      \draw [help lines] (-6,-6) grid (0,0);
      \draw [color=black,->] (-5.0,0) -- (5.0,0);
      \draw [color=black,->] (0,-5.0) -- (0,5.0);

      \draw [color=gray, line width=0.5pt, solid] (4,-3) circle (6.083);
      \draw [color=gray, line width=0.5pt, solid] (4,-3) circle (7.071);
      \draw [color=gray, line width=0.5pt, solid] (4,-3) circle (8.062);
      \draw [color=gray, line width=0.5pt, solid] (4,-3) circle (4.472);

      \draw[color=darkgray, line width=1.0pt, dashed, ->] (-1,2) -- (-4,-4);

      \filldraw [color=blue] (3,3) circle (2.5pt );
      \filldraw [color=blue] (-1,2) circle (2.5pt );
      \filldraw [color=white] (-1,2) circle (2.1pt );
      \filldraw [color=blue] (-4,-4) circle (2.5pt );
      \filldraw [color=blue] (0,-1) circle (2.5pt );
      \filldraw [color=blue] (4,-3) circle (2.5pt );
      \filldraw [color=yellow] (4,-3) circle (2.1pt );

      \filldraw [color=red] (2,1) circle (2.5pt );

      \node at (3.3,2.7) {$x_1$};
      \node at (-0.7,1.7) {$x_2$};
      \node at (-3.7,-4.3) {$x_3$};
      \node at (0.3,-1.3) {$x_4$};
      \node at (4.3,-3.3) {$x_5$};

      \node at (2.3,0.7) {$q$};
      \end{tikzpicture}
  \end{column}
 \end{columns}
\end{frame}

%Beispiel3 c:=x_2
\begin{frame}{Annulus Method}{Example}
 \begin{columns}[T]
  \begin{column}{.35\textwidth}
    \begin{itemize}
     \item Set $c:=x_2$ with $d(c, q) \approx 3.162$
     \item Set $s:=x_3$ with $d(s, q) \approx 7.810$
     \item $8.062 \approx d(p^*, s) > d(p^*, q) + d(c,q) \approx 7.634 \Rightarrow$ ignore items above $x_3$
    \end{itemize}
  \end{column}
  \begin{column}{.60\textwidth}
     \begin{tikzpicture}[%
      scale=0.6,
      ]
      \clip (-5.0,-5.0) rectangle (5.0,5.0);
      \draw [help lines] (0,0) grid (6,5);
      \draw [help lines] (-6,0) grid (0,5);
      \draw [help lines] (0,-6) grid (6,0);
      \draw [help lines] (-6,-6) grid (0,0);
      \draw [color=black,->] (-5.0,0) -- (5.0,0);
      \draw [color=black,->] (0,-5.0) -- (0,5.0);

      \draw [color=gray, line width=0.5pt, solid] (4,-3) circle (6.083);
      \draw [color=gray, line width=0.5pt, solid] (4,-3) circle (7.071);
      \draw [color=gray, line width=0.5pt, solid] (4,-3) circle (8.062);
      \draw [color=gray, line width=0.5pt, solid] (4,-3) circle (4.472);

      \draw[color=darkgray, line width=1.0pt, dashed, ->] (-1,2) -- (-4,-4);

      \draw[color=gray, line width=0.5pt, dashed] (4,-3) -- (-4,-4);
      \draw[color=gray, line width=0.5pt, dashed] (4,-3) -- (2,1);
      \draw[color=gray, line width=0.5pt, dashed] (-1,2) -- (2,1);

      \filldraw [color=blue] (3,3) circle (2.5pt );
      \filldraw [color=blue] (-1,2) circle (2.5pt );
      \filldraw [color=white] (-1,2) circle (2.1pt );
      \filldraw [color=blue] (-4,-4) circle (2.5pt );
      \filldraw [color=blue] (0,-1) circle (2.5pt );
      \filldraw [color=blue] (4,-3) circle (2.5pt );
      \filldraw [color=yellow] (4,-3) circle (2.1pt );

      \filldraw [color=red] (2,1) circle (2.5pt );

      \node at (3.3,2.7) {$x_1$};
      \node at (-0.7,1.7) {$x_2$};
      \node at (-3.7,-4.3) {$x_3$};
      \node at (0.3,-1.3) {$x_4$};
      \node at (4.3,-3.3) {$x_5$};

      \node at (2.3,0.7) {$q$};
      \end{tikzpicture}
  \end{column}
 \end{columns}
\end{frame}

%Beispiel3 c:=x_2
\begin{frame}<handout:0>{Annulus Method}{Example}
 \begin{columns}[T]
  \begin{column}{.35\textwidth}
    \begin{itemize}
     \item Set $s:=x_1$ with $d(s, q) \approx 2.236$
    \end{itemize}
  \end{column}
  \begin{column}{.60\textwidth}
     \begin{tikzpicture}[%
      scale=0.6,
      ]
      \clip (-5.0,-5.0) rectangle (5.0,5.0);
      \draw [help lines] (0,0) grid (6,5);
      \draw [help lines] (-6,0) grid (0,5);
      \draw [help lines] (0,-6) grid (6,0);
      \draw [help lines] (-6,-6) grid (0,0);
      \draw [color=black,->] (-5.0,0) -- (5.0,0);
      \draw [color=black,->] (0,-5.0) -- (0,5.0);

      \draw [color=gray, line width=0.5pt, solid] (4,-3) circle (6.083);
      \draw [color=gray, line width=0.5pt, solid] (4,-3) circle (7.071);
      \draw [color=gray, line width=0.5pt, solid] (4,-3) circle (8.062);
      \draw [color=gray, line width=0.5pt, solid] (4,-3) circle (4.472);

      \draw[color=darkgray, line width=1.0pt, dashed, ->] (-1,2) -- (-4,-4);
      \draw[color=darkgray, line width=1.0pt, dashed, ->] (-4,-4) -- (3,3);

      \filldraw [color=blue] (3,3) circle (2.5pt );
      \filldraw [color=blue] (-1,2) circle (2.5pt );
      \filldraw [color=white] (-1,2) circle (2.1pt );
      \filldraw [color=gray] (-4,-4) circle (2.5pt );
      \filldraw [color=blue] (0,-1) circle (2.5pt );
      \filldraw [color=blue] (4,-3) circle (2.5pt );
      \filldraw [color=yellow] (4,-3) circle (2.1pt );

      \filldraw [color=red] (2,1) circle (2.5pt );

      \node at (3.3,2.7) {$x_1$};
      \node at (-0.7,1.7) {$x_2$};
      \node at (-3.7,-4.3) {$x_3$};
      \node at (0.3,-1.3) {$x_4$};
      \node at (4.3,-3.3) {$x_5$};

      \node at (2.3,0.7) {$q$};
      \end{tikzpicture}
  \end{column}
 \end{columns}
\end{frame}

%Beispiel3 c:=x_2
\begin{frame}{Annulus Method}{Example}
 \begin{columns}[T]
  \begin{column}{.35\textwidth}
    \begin{itemize}
     \item Set $s:=x_1$ with $d(s, q) \approx 2.236$
     \item As $d(s, q) < d(c, q)$, set $c:=s$
    \end{itemize}
  \end{column}
  \begin{column}{.60\textwidth}
     \begin{tikzpicture}[%
      scale=0.6,
      ]
      \clip (-5.0,-5.0) rectangle (5.0,5.0);
      \draw [help lines] (0,0) grid (6,5);
      \draw [help lines] (-6,0) grid (0,5);
      \draw [help lines] (0,-6) grid (6,0);
      \draw [help lines] (-6,-6) grid (0,0);
      \draw [color=black,->] (-5.0,0) -- (5.0,0);
      \draw [color=black,->] (0,-5.0) -- (0,5.0);

      \draw [color=gray, line width=0.5pt, solid] (4,-3) circle (6.083);
      \draw [color=gray, line width=0.5pt, solid] (4,-3) circle (7.071);
      \draw [color=gray, line width=0.5pt, solid] (4,-3) circle (8.062);
      \draw [color=gray, line width=0.5pt, solid] (4,-3) circle (4.472);

      \draw[color=darkgray, line width=1.0pt, dashed, ->] (-1,2) -- (-4,-4);
      \draw[color=darkgray, line width=1.0pt, dashed, ->] (-4,-4) -- (3,3);

      \draw[color=gray, line width=0.5pt, dashed] (3,3) -- (2,1);
      \draw[color=gray, line width=0.5pt, dashed] (-1,2) -- (2,1);

      \filldraw [color=blue] (3,3) circle (2.5pt );
      \filldraw [color=blue] (-1,2) circle (2.5pt );
      \filldraw [color=white] (-1,2) circle (2.1pt );
      \filldraw [color=gray] (-4,-4) circle (2.5pt );
      \filldraw [color=blue] (0,-1) circle (2.5pt );
      \filldraw [color=blue] (4,-3) circle (2.5pt );
      \filldraw [color=yellow] (4,-3) circle (2.1pt );

      \filldraw [color=red] (2,1) circle (2.5pt );

      \node at (3.3,2.7) {$x_1$};
      \node at (-0.7,1.7) {$x_2$};
      \node at (-3.7,-4.3) {$x_3$};
      \node at (0.3,-1.3) {$x_4$};
      \node at (4.3,-3.3) {$x_5$};

      \node at (2.3,0.7) {$q$};
      \end{tikzpicture}
  \end{column}
 \end{columns}
\end{frame}

%Beispiel3 c:=x_1
\begin{frame}<handout:0>{Annulus Method}{Example}
 \begin{columns}[T]
  \begin{column}{.35\textwidth}
    \begin{itemize}
     \item Set $c:=x_1$ with $d(c, q) \approx 2.236$
    \end{itemize}
  \end{column}
  \begin{column}{.60\textwidth}
     \begin{tikzpicture}[%
      scale=0.6,
      ]
      \clip (-5.0,-5.0) rectangle (5.0,5.0);
      \draw [help lines] (0,0) grid (6,5);
      \draw [help lines] (-6,0) grid (0,5);
      \draw [help lines] (0,-6) grid (6,0);
      \draw [help lines] (-6,-6) grid (0,0);
      \draw [color=black,->] (-5.0,0) -- (5.0,0);
      \draw [color=black,->] (0,-5.0) -- (0,5.0);

      \draw [color=gray, line width=0.5pt, solid] (4,-3) circle (6.083);
      \draw [color=gray, line width=0.5pt, solid] (4,-3) circle (7.071);
      \draw [color=gray, line width=0.5pt, solid] (4,-3) circle (8.062);
      \draw [color=gray, line width=0.5pt, solid] (4,-3) circle (4.472);

      \filldraw [color=blue] (3,3) circle (2.5pt );
      \filldraw [color=white] (3,3) circle (2.1pt );
      \filldraw [color=gray] (-1,2) circle (2.5pt );
      \filldraw [color=gray] (-4,-4) circle (2.5pt );
      \filldraw [color=blue] (0,-1) circle (2.5pt );
      \filldraw [color=blue] (4,-3) circle (2.5pt );
      \filldraw [color=yellow] (4,-3) circle (2.1pt );

      \filldraw [color=red] (2,1) circle (2.5pt );

      \node at (3.3,2.7) {$x_1$};
      \node at (-0.7,1.7) {$x_2$};
      \node at (-3.7,-4.3) {$x_3$};
      \node at (0.3,-1.3) {$x_4$};
      \node at (4.3,-3.3) {$x_5$};

      \node at (2.3,0.7) {$q$};
      \end{tikzpicture}
  \end{column}
 \end{columns}
\end{frame}

%Beispiel3 c:=x_1
\begin{frame}<handout:0>{Annulus Method}{Example}
 \begin{columns}[T]
  \begin{column}{.35\textwidth}
    \begin{itemize}
     \item Set $c:=x_1$ with $d(c, q) \approx 2.236$
     \item Set $s:=x_4$ with $d(s, q) \approx 2.828$
    \end{itemize}
  \end{column}
  \begin{column}{.60\textwidth}
     \begin{tikzpicture}[%
      scale=0.6,
      ]
      \clip (-5.0,-5.0) rectangle (5.0,5.0);
      \draw [help lines] (0,0) grid (6,5);
      \draw [help lines] (-6,0) grid (0,5);
      \draw [help lines] (0,-6) grid (6,0);
      \draw [help lines] (-6,-6) grid (0,0);
      \draw [color=black,->] (-5.0,0) -- (5.0,0);
      \draw [color=black,->] (0,-5.0) -- (0,5.0);

      \draw [color=gray, line width=0.5pt, solid] (4,-3) circle (6.083);
      \draw [color=gray, line width=0.5pt, solid] (4,-3) circle (7.071);
      \draw [color=gray, line width=0.5pt, solid] (4,-3) circle (8.062);
      \draw [color=gray, line width=0.5pt, solid] (4,-3) circle (4.472);

      \draw[color=darkgray, line width=1.0pt, dashed, ->] (3,3) -- (0,-1);

      \filldraw [color=blue] (3,3) circle (2.5pt );
      \filldraw [color=white] (3,3) circle (2.1pt );
      \filldraw [color=gray] (-1,2) circle (2.5pt );
      \filldraw [color=gray] (-4,-4) circle (2.5pt );
      \filldraw [color=blue] (0,-1) circle (2.5pt );
      \filldraw [color=blue] (4,-3) circle (2.5pt );
      \filldraw [color=yellow] (4,-3) circle (2.1pt );

      \filldraw [color=red] (2,1) circle (2.5pt );

      \node at (3.3,2.7) {$x_1$};
      \node at (-0.7,1.7) {$x_2$};
      \node at (-3.7,-4.3) {$x_3$};
      \node at (0.3,-1.3) {$x_4$};
      \node at (4.3,-3.3) {$x_5$};

      \node at (2.3,0.7) {$q$};
      \end{tikzpicture}
  \end{column}
 \end{columns}
\end{frame}

%Beispiel3 c:=x_1
\begin{frame}{Annulus Method}{Example}
 \begin{columns}[T]
  \begin{column}{.35\textwidth}
    \begin{itemize}
     \item Set $c:=x_1$ with $d(c, q) \approx 2.236$
     \item Set $s:=x_4$ with $d(s, q) \approx 2.828$
     \item $4.472 \approx d(p^*, s) > d(p^*, q) - d(c,q) \approx 2.236 \Rightarrow$ ignore no items
    \end{itemize}
  \end{column}
  \begin{column}{.60\textwidth}
     \begin{tikzpicture}[%
      scale=0.6,
      ]
      \clip (-5.0,-5.0) rectangle (5.0,5.0);
      \draw [help lines] (0,0) grid (6,5);
      \draw [help lines] (-6,0) grid (0,5);
      \draw [help lines] (0,-6) grid (6,0);
      \draw [help lines] (-6,-6) grid (0,0);
      \draw [color=black,->] (-5.0,0) -- (5.0,0);
      \draw [color=black,->] (0,-5.0) -- (0,5.0);

      \draw [color=gray, line width=0.5pt, solid] (4,-3) circle (6.083);
      \draw [color=gray, line width=0.5pt, solid] (4,-3) circle (7.071);
      \draw [color=gray, line width=0.5pt, solid] (4,-3) circle (8.062);
      \draw [color=gray, line width=0.5pt, solid] (4,-3) circle (4.472);

      \draw[color=darkgray, line width=1.0pt, dashed, ->] (3,3) -- (0,-1);

      \draw[color=gray, line width=0.5pt, dashed] (4,-3) -- (0,-1);
      \draw[color=gray, line width=0.5pt, dashed] (4,-3) -- (2,1);
      \draw[color=gray, line width=0.5pt, dashed] (3,3) -- (2,1);

      \filldraw [color=blue] (3,3) circle (2.5pt );
      \filldraw [color=white] (3,3) circle (2.1pt );
      \filldraw [color=gray] (-1,2) circle (2.5pt );
      \filldraw [color=gray] (-4,-4) circle (2.5pt );
      \filldraw [color=blue] (0,-1) circle (2.5pt );
      \filldraw [color=blue] (4,-3) circle (2.5pt );
      \filldraw [color=yellow] (4,-3) circle (2.1pt );

      \filldraw [color=red] (2,1) circle (2.5pt );

      \node at (3.3,2.7) {$x_1$};
      \node at (-0.7,1.7) {$x_2$};
      \node at (-3.7,-4.3) {$x_3$};
      \node at (0.3,-1.3) {$x_4$};
      \node at (4.3,-3.3) {$x_5$};

      \node at (2.3,0.7) {$q$};
      \end{tikzpicture}
  \end{column}
 \end{columns}
\end{frame}

%Beispiel3 c:=x_1
\begin{frame}<handout:0>{Annulus Method}{Example}
 \begin{columns}[T]
  \begin{column}{.35\textwidth}
    \begin{itemize}
     \item Set $s:=x_5$ with $d(s, q) \approx 4.472 < 2.236 \approx d(c, q)$
    \end{itemize}
  \end{column}
  \begin{column}{.60\textwidth}
     \begin{tikzpicture}[%
      scale=0.6,
      ]
      \clip (-5.0,-5.0) rectangle (5.0,5.0);
      \draw [help lines] (0,0) grid (6,5);
      \draw [help lines] (-6,0) grid (0,5);
      \draw [help lines] (0,-6) grid (6,0);
      \draw [help lines] (-6,-6) grid (0,0);
      \draw [color=black,->] (-5.0,0) -- (5.0,0);
      \draw [color=black,->] (0,-5.0) -- (0,5.0);

      \draw [color=gray, line width=0.5pt, solid] (4,-3) circle (6.083);
      \draw [color=gray, line width=0.5pt, solid] (4,-3) circle (7.071);
      \draw [color=gray, line width=0.5pt, solid] (4,-3) circle (8.062);
      \draw [color=gray, line width=0.5pt, solid] (4,-3) circle (4.472);

      \draw[color=darkgray, line width=1.0pt, dashed, ->] (3,3) -- (0,-1);
      \draw[color=darkgray, line width=1.0pt, dashed, ->] (0,-1) -- (4,-3);

      \filldraw [color=blue] (3,3) circle (2.5pt );
      \filldraw [color=white] (3,3) circle (2.1pt );
      \filldraw [color=gray] (-1,2) circle (2.5pt );
      \filldraw [color=gray] (-4,-4) circle (2.5pt );
      \filldraw [color=gray] (0,-1) circle (2.5pt );
      \filldraw [color=blue] (4,-3) circle (2.5pt );
      \filldraw [color=yellow] (4,-3) circle (2.1pt );

      \filldraw [color=red] (2,1) circle (2.5pt );

      \node at (3.3,2.7) {$x_1$};
      \node at (-0.7,1.7) {$x_2$};
      \node at (-3.7,-4.3) {$x_3$};
      \node at (0.3,-1.3) {$x_4$};
      \node at (4.3,-3.3) {$x_5$};

      \node at (2.3,0.7) {$q$};
      \end{tikzpicture}
  \end{column}
 \end{columns}
\end{frame}

%Beispiel3 c:=x_1
\begin{frame}{Annulus Method}{Example}
 \begin{columns}[T]
  \begin{column}{.35\textwidth}
    \begin{itemize}
     \item Set $s:=x_5$ with $d(s, q) \approx 4.472 < 2.236 \approx d(c, q)$
     \item End of list, $c = x_1$ nearest neighbour of $q$
    \end{itemize}
  \end{column}
  \begin{column}{.60\textwidth}
     \begin{tikzpicture}[%
      scale=0.6,
      ]
      \clip (-5.0,-5.0) rectangle (5.0,5.0);
      \draw [help lines] (0,0) grid (6,5);
      \draw [help lines] (-6,0) grid (0,5);
      \draw [help lines] (0,-6) grid (6,0);
      \draw [help lines] (-6,-6) grid (0,0);
      \draw [color=black,->] (-5.0,0) -- (5.0,0);
      \draw [color=black,->] (0,-5.0) -- (0,5.0);

      \draw [color=gray, line width=0.5pt, solid] (4,-3) circle (6.083);
      \draw [color=gray, line width=0.5pt, solid] (4,-3) circle (7.071);
      \draw [color=gray, line width=0.5pt, solid] (4,-3) circle (8.062);
      \draw [color=gray, line width=0.5pt, solid] (4,-3) circle (4.472);

      \draw[color=darkgray, line width=1.0pt, dashed, ->] (3,3) -- (0,-1);
      \draw[color=darkgray, line width=1.0pt, dashed, ->] (0,-1) -- (4,-3);

      \filldraw [color=blue] (3,3) circle (2.5pt );
      \filldraw [color=white] (3,3) circle (2.1pt );
      \filldraw [color=gray] (-1,2) circle (2.5pt );
      \filldraw [color=gray] (-4,-4) circle (2.5pt );
      \filldraw [color=gray] (0,-1) circle (2.5pt );
      \filldraw [color=blue] (4,-3) circle (2.5pt );
      \filldraw [color=yellow] (4,-3) circle (2.1pt );

      \filldraw [color=red] (2,1) circle (2.5pt );

      \node at (3.3,2.7) {$x_1$};
      \node at (-0.7,1.7) {$x_2$};
      \node at (-3.7,-4.3) {$x_3$};
      \node at (0.3,-1.3) {$x_4$};
      \node at (4.3,-3.3) {$x_5$};

      \node at (2.3,0.7) {$q$};
      \end{tikzpicture}
  \end{column}
 \end{columns}
\end{frame}

% Vor- und Nachteile
\begin{frame}{Annulus Method}{Advantages and disadvantages}
 \begin{block}{Advantages}
  \begin{itemize}
   \item Faster as Full Search
   \item Less memory usage than Orchards Algorithm
  \end{itemize}
 \end{block}
\end{frame}

\subsection{AESA}
\begin{frame}{AESA}{Approximating and Eliminating Search Algorithm}
 \begin{itemize}
  \item Create matrix with all distances $d(x, y)$, with $x, y \in S$
  \pause
  \item Every item is always in one status
  \begin{itemize}
   \item \textit{Known}, $d(x, q)$ is known
   \pause
   \item \textit{Unknown}, only $d_P(x, q)$ is known
   \pause
   \item \textit{Rejected}, $d_P(x, q)$ is bigger as smallest known distance $r$
  \end{itemize}
  \pause
  \item All $x \in S$ are \textit{Unknown} and $d_P(x, q) = -\infty$
  \pause
  \item Repeat until all $x \in S$ \textit{Known} oder \textit{Rejected}
  \begin{enumerate}
   \item Choose \textit{Unknown} item $x \in S$ with smallest $d_P(x, q)$
   \pause
   \item Calculate $d(x,q)$, so that $x$ gets \textit{Known}
   \pause
   \item Refresh the smallest known distance $r$
   \pause
   \item Set $P:= P \cup \{x\}$, refresh $d_P(x', q)$, if $x'$ is \textit{Unknown} mark $x'$ as \textit{Rejected}, if
    $d_P(x', q) > r$
  \end{enumerate}
 \end{itemize}
\end{frame}

\begin{frame}{LAESA}{\textit{Linear} Approximating and Eliminating Search Algorithm}
 \begin{itemize}
  \item Works with a set of \textit{pivot} items instead of a matrix
  \pause
  \item Works best if \textit{pivot} items are strongly separated
 \end{itemize}
\end{frame}

\section{Outlook}
\begin{frame}{Outlook}
 \begin{itemize}
  \item Metric trees
  \item \dots
 \end{itemize}
\end{frame}

\section{Resources}
\begin{frame}
 \frametitle{Resources}
 \begin{itemize}
  \item Otto Forster, 2006, \textit{Analysis 2}, Friedr. Vieweg \& Sohn Verlag
  \item Kenneth L. Clarkson, 2005, \textit{Nearest-Neighbor Searching and Metric Space Dimensions}, http://kenclarkson.org/nn\_survey/p.pdf
  \end{itemize}
\end{frame}

\end{document}